---
title: "Sacred Values Signal Trustworthiness --- Reproducibility Pipeline (Study 3)"
subtitle: "The incentivized trust game: behaviourally real and warranted trust"
author: "Sacred Values project"
date: "Knitted `r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    number_sections: true
    code_folding: show
    df_print: kable
    highlight: tango
---

<!-- Rscript -e 'rmarkdown::render("scripts/R/reproducibility_study_3.Rmd")' -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.align = "center", fig.width = 7.5, fig.height = 4.2)
```

# What this document is

Study 3 is the **incentivized lab trust game** (MonLEE, Monash; international pool,
N = 242). It is the behavioural anchor of the paper: it tests whether perceived
trustworthiness translates into *trusting behaviour with real money*, and whether that
trust is *warranted* --- whether sacred valuers actually return more when they could defect.

This pipeline reproduces, from the archived data, **every statistical test reported for
Study 3 --- in both the main text and the Supporting Information (SI)** --- and audits each
computed value against the number the *current manuscript* prints. There is a single source
of truth: the present main text and SI.

**Main-text coverage.** (i) The warranted-trustworthiness result --- the Pearson correlation
and penalised GAM of the proportion returned on sacred values *held* (`SacPers`), the paper's
single-panel **Figure 2**, reported as +0.31 percentage points per value held. (ii) The
now-primary investor result: the **categorical (binary) contrast** between trustees who
revealed at least one sacred value and those who revealed only tradeable ones (SI
`sec:s3-binary`), including its controls, its head-to-head against the non-sacred count, and
its convergence on stated beliefs. (iii) A short check that the analogous **trustee-side
binary is null** --- the warranted-trust side is genuinely graded, not categorical --- which
backs the main text's asymmetry claim.

**SI coverage.** The continuous net-balance fit and its controls (`sec:s3-continuous`); the
alternative regression specifications (`s3-alt-specs`); the post-game **belief** correlations
(`s3-scatter-belief`, `s3-scatter-trustworthiness`); the willingness-to-revise correlation
(`s3-survey-revision`); the **revelation-preference** analysis --- the preregistered
chi-square and the conservative mixed-effects logistic regression (`s3-mixed-logit`); the
**trust-connoting content** bake-off (`s3-trust-index`) and its factor-loading companion; and
the **religiosity** rank-sum test and table (`s3-religiosity`).

The trustworthiness predictor is `SacPers` (sacred values *held*) by project decision --- the
conceptually correct measure for "do people who hold sacred values behave more trustworthily?"
and the one plotted in Figure 2. It follows the same push-button standard as the Studies 1--2
and 4--5 pipelines (self-contained, data auto-located, targets declared in advance,
`sessionInfo()` footer).

**Provenance note.** The OSF archive ships the *preregistered* Stata analysis
(`hypothesis 1/2/3 *.do`: OLS, rank-sum, mixed logit). The paper departs from the
preregistered investor specification (the raw revealed count) in favour of the categorical
test the matured index-signal account implies; both the preregistered count model and the
categorical test are reported. The preregistered OLS port is included in an appendix below for
transparency. The GAM/correlation R code itself was never archived, so those analyses are
rebuilt here from the specifications the manuscript describes.

---

# Setup

```{r packages}
suppressPackageStartupMessages({
  library(haven)    # read Stata .dta
  library(dplyr)    # data manipulation
  library(mgcv)     # generalized additive models
  library(lme4)     # mixed logistic regression (Stata melogit equivalent)
  library(coin)     # Mann-Whitney with the exact standardized Z
  library(ggplot2)  # figures
  library(tibble)
  library(knitr)
})
set.seed(20260619)
```

```{r data-root}
find_data_root <- function(start = getwd(), marker = "osf") {
  d <- normalizePath(start, winslash = "/", mustWork = FALSE)
  repeat {
    if (dir.exists(file.path(d, marker))) return(d)
    parent <- dirname(d)
    if (identical(parent, d))
      stop("Could not locate the 'osf' data directory from ", start, call. = FALSE)
    d <- parent
  }
}

S3_DIR <- file.path(find_data_root(), "osf", "Study 3 - Lab Study",
                    "Replication Documentation", "Data", "Analysis Data")
s3_file <- file.path(S3_DIR, "combined-ult-trust-withfactors.dta")
stopifnot("Study 3 analysis .dta not found" = file.exists(s3_file))
cat("Study 3 data:\n  ", s3_file, "\n")
```

## Helper functions

```{r helpers}
# Pearson correlation, reported the way the paper does ---------------------------
pearson_report <- function(x, y) {
  ok <- stats::complete.cases(x, y)
  ct <- suppressWarnings(stats::cor.test(x[ok], y[ok], method = "pearson"))
  list(r = unname(ct$estimate), p = ct$p.value, n = sum(ok))
}

# Common-language effect size ----------------------------------------------------
# P(a random value from the group coded 1 exceeds a random value from group 0),
# ties counted as one half -- the same statistic the vignette studies report.
cles <- function(x, g) {
  a <- x[g == 1 & !is.na(x)]; b <- x[g == 0 & !is.na(x)]
  mean(outer(a, b, function(u, v) (u > v) + 0.5 * (u == v)))
}

# Pull the smooth-term F, p, edf from a fitted mgcv::gam ------------------------
# edf (effective degrees of freedom) is the diagnostic for non-linearity: edf ~ 1
# means the smooth is effectively a straight line (the GAM only *relaxed* linearity,
# it did not find curvature); edf materially > 1 means genuine curvature.
gam_smooth <- function(fit) {
  s  <- summary(fit)
  st <- s$s.table[1, ]
  list(F = unname(st["F"]), p = unname(st["p-value"]), edf = unname(st["edf"]),
       n = s$n, r2 = s$r.sq)
}

# Reported one-SD effect: predicted change in the outcome for a one-SD increase in the
# focal predictor, every covariate held at its mean. This is the "B" the manuscript
# reports (e.g. the +8 pp/SD trust effect). ------------------------------------------
gam_oneSD <- function(fit, xvar) {
  md    <- fit$model                       # model frame: response first, then predictors
  preds <- setdiff(names(md), names(md)[1])
  mk <- function(xval) as.data.frame(setNames(
    lapply(preds, function(p) if (p == xvar) xval else mean(md[[p]], na.rm = TRUE)), preds))
  mu <- mean(md[[xvar]], na.rm = TRUE); s <- sd(md[[xvar]], na.rm = TRUE)
  unname(predict(fit, mk(mu + s)) - predict(fit, mk(mu)))
}

# One audit row: computed vs the value the manuscript reports ------------------
audit_row <- function(quantity, computed, claimed, tol, digits = 2) {
  fmt <- function(v) if (is.na(v)) "--" else formatC(v, format = "f", digits = digits)
  check <- if (is.na(claimed)) "descriptive"
           else { d <- abs(computed - claimed)
                  if (d <= tol) "match" else if (d <= 3 * tol) "close" else "DIFFERS" }
  tibble(Quantity = quantity, Reported = fmt(claimed), Computed = fmt(computed), Check = check)
}
```

## Load and prepare

```{r prep}
raw <- read_dta(s3_file)

# --- Revealed-similarity ("self-similarity") factors --------------------------------
# Reconstructs `hypothesis 2 code.do`: min-max normalise the three sacred-inventory
# factor scores (full inventory and revealed-only), weight the revealed scores by the
# number of statements the trustee revealed, then for each trustee take the negated
# Euclidean distance between the investor partner's full-inventory score and the
# trustee's weighted revealed score. Higher `wsim` = investor and trustee look more
# similar on that factor. Built **once** here so both the controlled models and the
# preregistered OLS port (H2, appendix) consume a single construction.
nrm <- function(x) { r <- range(x, na.rm = TRUE); (x - r[1]) / (r[2] - r[1]) }
full <- raw                                   # normalise over the full archive, per the .do order
for (i in 1:3) {
  full[[paste0("f", i, "n")]]  <- nrm(full[[paste0("full_f", i)]])
  full[[paste0("r", i, "n")]]  <- nrm(full[[paste0("rev_f", i)]])
  full[[paste0("r", i, "n2")]] <- full[[paste0("r", i, "n")]] / full$NumRev
}
full <- full %>% filter(as_factor(treatment) == "trust") %>%   # trust game only (drop ultimatum arm)
  group_by(t3groups) %>%
  mutate(pf1 = sum(f1n * (role_a_b == 1), na.rm = TRUE),        # investor partner's factor score
         pf2 = sum(f2n * (role_a_b == 1), na.rm = TRUE),
         pf3 = sum(f3n * (role_a_b == 1), na.rm = TRUE)) %>%
  ungroup()
for (i in 1:3)                                                 # revealed similarity = -|partner - revealed|
  full[[paste0("wsim", i)]] <- -sqrt((full[[paste0("pf", i)]] - full[[paste0("r", i, "n2")]])^2)

# Demographic controls (per the OSF `.do` file notes): gender 1 = men, age in years.
# NB these sit on the *trustee* row in the controlled models, matching the .do files'
# `if role_a_b == 2` regressions. An earlier write-up described these controls as the
# investor's; the archived code uses the trustee's. Flagged in the audit.
full <- full %>%
  mutate(men = as.integer(as_factor(survey1playersex) == "Male"),
         age = as.numeric(survey1playerage))

# Roles: 1 = investor, 2 = trustee (per the OSF .do files).
investors <- full %>% filter(role_a_b == 1)
trustees  <- full %>% filter(role_a_b == 2)

cat(sprintf("Trust-game rows: %d  (investors %d, trustees %d = %d complete dyads)\n",
            nrow(full), nrow(investors), nrow(trustees), nrow(trustees)))
cat("Archived sample = 242 participants = 121 complete dyads; matches the manuscript N = 242.\n")
# Confirm the dyads are intact (equal investor / trustee counts).
stopifnot("unexpected role imbalance (broken dyads)" = nrow(investors) == nrow(trustees))
```

Key variables (from the OSF data and `.do` files):

- `Tru_diff` --- investor trust: change in amount transferred between Trust Games 2 and 1.
- `Wor_agg_avg` --- trustworthiness: average proportion returned across Games 1 and 2.
- `SacPers` --- sacred values *held*: endorsed in the Phase-1 inventory.
- `SacRev` --- sacred values *revealed* (count) by the trustee.
- `NetRev` = `SacRev - NonSacRev` --- net (balance of) sacred values revealed by the trustee.
- `partnerNetRev` --- the trustee partner's `NetRev`, as seen by the investor.
- `wsim1/wsim2/wsim3` --- revealed self-similarity on three factors (built in prep; see above).
- `Greed`, `age` (`survey1playerage`), `men` (`survey1playersex`) --- controls.

---

# Main-text results

## Warranted trustworthiness (Figure 2)

The warranted-trust claim: trustees who *hold* more sacred values (`SacPers`, the Phase-2
inventory) return a greater share of the stake. This is the paper's single-panel **Figure 2**
and the `SacPers` correlation. The continuous investor association (net balance revealed) is
computed here too, but it is an SI result (`sec:s3-continuous`) --- the *primary* investor
result is the categorical test in the next section.

```{r pearson}
# Warranted trust (Fig 2): trustees who hold more sacred values return more.
s3_return <- pearson_report(trustees$Wor_agg_avg, trustees$SacPers)
# Continuous investor association (net balance revealed) -- an SI result, kept here as a
# descriptive cross-reference to the primary categorical test below.
s3_invest <- pearson_report(investors$Tru_diff, investors$partnerNetRev)

cat(sprintf("Fig 2  trustworthiness ~ sacred values held  : r = %.3f, p = %.5f (n = %d)\n",
            s3_return$r, s3_return$p, s3_return$n))
cat(sprintf("(SI)   investor trust ~ net values revealed  : r = %.3f, p = %.4f (n = %d)\n",
            s3_invest$r, s3_invest$p, s3_invest$n))
```

**Percentage-point interpretation.** The manuscript reports the trustworthiness effect as
**+0.31 percentage points** of the stake returned per sacred value held, and **+5.7 points**
across a one-standard-deviation increase. `Wor_agg_avg` is the proportion returned, so the
linear `SacPers` slope times 100 is the pp figure.

```{r maintext-worthy}
b_sp  <- coef(lm(Wor_agg_avg ~ SacPers, data = trustees))["SacPers"]
sd_sp <- sd(trustees$SacPers, na.rm = TRUE)
cat(sprintf("Trustworthiness: +%.2f pp per sacred value held; +%.1f pp per 1 SD (%.1f values); p = %.5f.\n",
            100 * b_sp, 100 * b_sp * sd_sp, sd_sp,
            summary(lm(Wor_agg_avg ~ SacPers, data = trustees))$coefficients["SacPers", "Pr(>|t|)"]))
```

The reproduction figure is a **single trustworthiness panel** matching the current manuscript
Figure 2: the penalised GAM (k = 3) of proportion returned on sacred values held, raw scatter
removed and the vertical axis zoomed to the fitted curve and its 95% band, so the band's
departure from the flat no-effect line is legible. The manuscript PDF is written by the bespoke
script `scripts/R/study3_partials_figure.R` (`Figures/study3_trustgame_partials.pdf`); this
reproduces the same smooth.

```{r figure, fig.width=5.5, fig.height=4.2, fig.cap="Study 3, Figure 2 (single panel). Penalised GAM fit (k = 3, REML) of the proportion returned on the trustee's own sacred values held, with the raw scatter removed and the vertical axis zoomed. Line is the GAM fit; shaded band is the 95% CI; dashed line is the no-effect reference; tick marks show observed values held."}
# Manuscript Figure 2: stiff (k = 3) GAM partial fit of trustworthiness on values held,
# scatter removed and axis zoomed. Same smooth-only model as the trustworthiness GAM below.
gfig_worthy <- gam(Wor_agg_avg ~ s(SacPers, k = 3), data = trustees, method = "REML")

band_fig <- function(fit, xvar, n = 200) {
  x  <- fit$model[[xvar]]
  gx <- seq(min(x, na.rm = TRUE), max(x, na.rm = TRUE), length.out = n)
  pr <- predict(fit, setNames(data.frame(gx), xvar), se.fit = TRUE)
  data.frame(x = gx, fit = pr$fit,
             lwr = pr$fit - 1.96 * pr$se.fit, upr = pr$fit + 1.96 * pr$se.fit)
}
fig_df <- band_fig(gfig_worthy, "SacPers")
rug_df <- data.frame(x = gfig_worthy$model$SacPers)

ggplot(fig_df, aes(x, fit)) +
  geom_hline(yintercept = mean(fig_df$fit), linetype = "dashed",
             colour = "grey60", linewidth = 0.4) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "#1a73e8", alpha = 0.18) +
  geom_line(colour = "#1a73e8", linewidth = 1) +
  geom_rug(data = rug_df, aes(x = x), inherit.aes = FALSE, sides = "b",
           colour = "#1a73e8", alpha = 0.45) +
  scale_y_continuous(expand = expansion(mult = 0.12)) +
  labs(x = "Sacred values held (SacPers)", y = "Proportion returned (fitted)") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank())
```

## The investor signal: a categorical (binary) test

The now-primary investor result. On the index-signal account a revealed sacred value discloses
a *capacity* --- the ability to internalise a moral rule at all --- so the diagnostic contrast is
categorical: did the trustee reveal **any** sacred value (`SacRev > 0`), or only tradeable ones?
Every trustee revealed at least two statements, so the `SacRev == 0` group genuinely revealed
only values they had priced. This reproduces the main-text paragraph and SI `sec:s3-binary`;
the source of these numbers is `explorations/study3_netrev_decomposition/` and the figure script
`scripts/R/study3_investor_binary_violin.R`.

```{r binary-investor}
trustees <- trustees %>% mutate(any_sacred = as.integer(SacRev > 0))
n_trad <- sum(trustees$any_sacred == 0)   # revealed only tradeable values
n_sac  <- sum(trustees$any_sacred == 1)   # revealed >= 1 sacred value

# Group means of the change in trust (Tru_diff).
bin_means <- tapply(trustees$Tru_diff, trustees$any_sacred, mean, na.rm = TRUE)

# Headline contrast.
m_bin    <- lm(Tru_diff ~ any_sacred, data = trustees)
bin_b    <- coef(m_bin)["any_sacred"]
bin_p    <- summary(m_bin)$coefficients["any_sacred", "Pr(>|t|)"]
bin_r2   <- summary(m_bin)$r.squared
bin_cles <- cles(trustees$Tru_diff, trustees$any_sacred)

# Robust to controls (revealed self-similarity + dispositional greed; no age/sex,
# which have no theoretical warrant here).
m_binc  <- lm(Tru_diff ~ any_sacred + wsim1 + wsim2 + wsim3 + Greed, data = trustees)
binc_b  <- coef(m_binc)["any_sacred"]
binc_p  <- summary(m_binc)$coefficients["any_sacred", "Pr(>|t|)"]

# Genuine sacred threshold, not the non-sacred penalty in disguise: hold the non-sacred
# count (and total revealed) fixed and the threshold still moves trust.
m_binn  <- lm(Tru_diff ~ any_sacred + NonSacRev, data = trustees)
binn_b  <- coef(m_binn)["any_sacred"]
binn_p  <- summary(m_binn)$coefficients["any_sacred", "Pr(>|t|)"]
binn_np <- summary(m_binn)$coefficients["NonSacRev",  "Pr(>|t|)"]
m_binnn <- lm(Tru_diff ~ any_sacred + NonSacRev + NumRev, data = trustees)
binnn_b <- coef(m_binnn)["any_sacred"]
binnn_p <- summary(m_binnn)$coefficients["any_sacred", "Pr(>|t|)"]

# Convergence on stated beliefs: same split predicts the trustee's belief the partner
# will trust them more.
m_binbel <- lm(as.numeric(blf_trustee_trust) ~ any_sacred, data = trustees)
binbel_b <- coef(m_binbel)["any_sacred"]
binbel_p <- summary(m_binbel)$coefficients["any_sacred", "Pr(>|t|)"]

cat(sprintf("Groups: only-tradeable n = %d (mean Tru_diff %.3f) | some-sacred n = %d (mean %.3f)\n",
            n_trad, bin_means["0"], n_sac, bin_means["1"]))
cat(sprintf("Binary contrast   : b = %.3f, p = %.4f, CLES = %.3f, R^2 = %.3f\n",
            bin_b, bin_p, bin_cles, bin_r2))
cat(sprintf("  + controls      : b = %.3f, p = %.4f  (self-similarity + greed)\n", binc_b, binc_p))
cat(sprintf("  + NonSacRev     : b = %.3f, p = %.4f  (NonSacRev p = %.3f, null)\n", binn_b, binn_p, binn_np))
cat(sprintf("  + NonSacRev+Num : b = %.3f, p = %.4f\n", binnn_b, binnn_p))
cat(sprintf("On belief 'partner trusts more' : b = %.3f, p = %.6f\n", binbel_b, binbel_p))
```

**Reading the result.** The split is well balanced (56 only-tradeable, 65 some-sacred). Trust
held roughly steady toward the some-sacred group (mean change ≈ -0.03) and was withdrawn from
the only-tradeable group (≈ -0.22) --- a gap of about 20 percentage points of the endowment
(b = .196, p = .002, CLES = .65), accounting for about 8% of the variance. It survives controls
for self-similarity and greed (b = .167, p = .01) where the continuous `NetRev` fit decays to
marginal, and it is a genuine sacred threshold rather than the non-sacred penalty in disguise:
holding the non-sacred count fixed, crossing the sacred threshold still raises trust (p = .028)
while the non-sacred count goes null. The same split predicts the trustee's belief that the
partner will trust them more.

## The trustee side is graded, not categorical

The mirror check the main text cites: binarising the trustworthiness predictor into *any* sacred
value held versus none does **not** on its own predict trustworthiness --- the warranted-trust
side is genuinely dose-response, and the continuous count subsumes the binary. This is why the
paper reports the investor side categorically and the trustee side continuously.

```{r binary-trustee}
trustees <- trustees %>% mutate(any_held = as.integer(SacPers > 0))
n_held0 <- sum(trustees$any_held == 0)   # holds no sacred value
n_held1 <- sum(trustees$any_held == 1)   # holds >= 1

tb      <- lm(Wor_agg_avg ~ any_held, data = trustees)
tb_b    <- coef(tb)["any_held"]
tb_p    <- summary(tb)$coefficients["any_held", "Pr(>|t|)"]
tb_cles <- cles(trustees$Wor_agg_avg, trustees$any_held)
tbc_p   <- summary(lm(Wor_agg_avg ~ any_held + Greed, data = trustees))$coefficients["any_held", "Pr(>|t|)"]

# Head-to-head: continuous count against the binary. Count survives, binary collapses.
tb_h2h   <- lm(Wor_agg_avg ~ any_held + SacPers, data = trustees)
tb_h2h_p <- summary(tb_h2h)$coefficients["any_held", "Pr(>|t|)"]
tb_sp_p  <- summary(tb_h2h)$coefficients["SacPers",  "Pr(>|t|)"]
add_sp_F <- anova(lm(Wor_agg_avg ~ any_held, data = trustees), tb_h2h)$F[2]  # adding count to binary
add_ah_F <- anova(lm(Wor_agg_avg ~ SacPers,  data = trustees), tb_h2h)$F[2]  # adding binary to count

# Median split (SI sec:s3-binary): the zero cut is low-powered because the held-none group is
# small (31) and the signal is spread across the whole 0-61 range. Cutting at the MEDIAN values
# held instead yields balanced groups and recovers the graded effect -- but it is still just a
# coarse proxy for the continuous count, not evidence of a categorical threshold.
med_held <- median(trustees$SacPers, na.rm = TRUE)
trustees <- trustees %>% mutate(hi_held = as.integer(SacPers >= med_held))
n_lo <- sum(trustees$hi_held == 0); n_hi <- sum(trustees$hi_held == 1)
mb      <- lm(Wor_agg_avg ~ hi_held, data = trustees)
mb_b    <- coef(mb)["hi_held"]
mb_p    <- summary(mb)$coefficients["hi_held", "Pr(>|t|)"]
mb_cles <- cles(trustees$Wor_agg_avg, trustees$hi_held)
mb_r2   <- summary(mb)$r.squared
mbc_p   <- summary(lm(Wor_agg_avg ~ hi_held + Greed, data = trustees))$coefficients["hi_held", "Pr(>|t|)"]
# Head-to-head: median split against the continuous count -- neither dominates (both proxy one gradient).
mb_h2h   <- lm(Wor_agg_avg ~ hi_held + SacPers, data = trustees)
mb_h2h_p <- summary(mb_h2h)$coefficients["hi_held", "Pr(>|t|)"]
mb_sp_p  <- summary(mb_h2h)$coefficients["SacPers", "Pr(>|t|)"]

cat(sprintf("Held none n = %d | holds >=1 n = %d\n", n_held0, n_held1))
cat(sprintf("Binary held>0 : b = %.3f (%.1f pp), p = %.3f, CLES = %.3f; + greed p = %.3f\n",
            tb_b, 100 * tb_b, tb_p, tb_cles, tbc_p))
cat(sprintf("Head-to-head  : any_held p = %.3f (null), SacPers p = %.4f (survives)\n", tb_h2h_p, tb_sp_p))
cat(sprintf("Nested F      : adding count to binary F = %.2f; adding binary to count F = %.2f\n",
            add_sp_F, add_ah_F))
cat(sprintf("Median split (>= %g held): low n = %d | high n = %d\n", med_held, n_lo, n_hi))
cat(sprintf("Median split  : b = %.3f (%.1f pp), p = %.4f, CLES = %.3f, R^2 = %.3f; + greed p = %.4f\n",
            mb_b, 100 * mb_b, mb_p, mb_cles, mb_r2, mbc_p))
cat(sprintf("  h2h vs count: median-split p = %.3f, SacPers p = %.3f (neither dominates; one gradient)\n",
            mb_h2h_p, mb_sp_p))
```

**Reading the result.** Dichotomising the trustee side *at zero* is marginal alone (+6.4 pp, CLES = .59,
p = .06), null under controls (p = .15), and the continuous count fully subsumes it (adding the
count to the binary: F ≈ 12, p < .001; adding the binary to the count: F ≈ 0.1, p = .73). The
trustee side is genuinely graded --- the reception-vs-trait asymmetry the manuscript describes.
The zero-cut null is a matter of power, not of the underlying relationship: only 31 of 121 trustees
held no sacred value, so the cut isolates a small, atypical group and discards the gradient in
between. Cutting at the **median** (11 values held) instead gives balanced groups (60 vs 61) and
recovers a clear contrast (+10.9 pp, p < .001, CLES = .68, R² = .11) that survives the greed control
(p = .002). But that is not a categorical threshold: entered against the continuous count neither
dominates (median split p = .37, count p = .16), so the median contrast is a coarse proxy for the
one underlying gradient. This backs the main-text point that the zero split loses power while a
balanced split recovers the effect.

# SI results

## Trustworthiness holds within each trust game (`sec:s3-pergame`)

Figure 2 regresses the trustee's return **averaged across both trust games** --- `Wor_agg_avg`
is the mean of `Wor_pre_avg` (the Phase-1 baseline game, played before the sacred-value
elicitation) and `Wor_post_avg` (the Phase-3 game, after the trustee's partial revelation).
Averaging is a noise-reduction move: the two games are two reads on one standing disposition
(they correlate at *r* ≈ .77), so the mean is the more reliable estimate. It is not manufacturing
the effect --- the dose-response is present in each game on its own, at similar magnitude. The
cleanest case is the Phase-1 baseline, played *before* the elicitation even occurred: values held
still predict returns there, so this is a standing trait, not reactivity to having just priced
one's values or consistency toward a partner who saw the revelation.

```{r pergame-worthy}
# Per-game components of Wor_agg_avg: Wor_pre_avg = Game 1 (Phase-1 baseline),
# Wor_post_avg = Game 2 (Phase-3, after revelation). Confirm the pooled measure is their mean.
stopifnot("Wor_agg_avg is not the mean of the two per-game measures" =
          max(abs(trustees$Wor_agg_avg -
                  (trustees$Wor_pre_avg + trustees$Wor_post_avg) / 2), na.rm = TRUE) < 1e-6)
rel <- cor(trustees$Wor_pre_avg, trustees$Wor_post_avg, use = "complete.obs")

# Linear + GAM report of a per-game outcome on SacPers (values held).
pg <- function(dv) {
  ok  <- stats::complete.cases(trustees[[dv]], trustees$SacPers)
  d   <- data.frame(y = trustees[[dv]][ok], SacPers = trustees$SacPers[ok])
  lm1 <- lm(y ~ SacPers, data = d)
  b   <- coef(lm1)["SacPers"]
  g   <- gam(y ~ s(SacPers, k = 3), data = d, method = "REML")
  c(r    = unname(cor(d$y, d$SacPers)),
    pp   = unname(100 * b),
    ppSD = unname(100 * b * sd(d$SacPers)),
    p    = summary(lm1)$coefficients["SacPers", "Pr(>|t|)"],
    R2   = summary(lm1)$r.squared,
    edf  = unname(summary(g)$s.table[1, "edf"]))
}
pg_tab <- rbind(`Both games (Fig 2)`  = pg("Wor_agg_avg"),
                `Game 1 (baseline)`   = pg("Wor_pre_avg"),
                `Game 2 (revelation)` = pg("Wor_post_avg"))

cat(sprintf("Between-game reliability: cor(Game 1, Game 2 return) = %.3f\n\n", rel))
print(round(pg_tab, 3))
```

**Reading the result.** The slope is positive, significant, and near-identical in each game
(Game 1: +0.28 pp/value, *p* = .001; Game 2: +0.35 pp/value, *p* < .001), and the pooled fit
(+0.31 pp/value, *p* < .001) sits between them with the tightest *p* --- exactly how averaging two
correlated measures should behave. Every smooth is linear (edf = 1.00). The averaging in Figure 2
is a reliability convenience, not a load-bearing analytic choice.

## Continuous net-balance fit and controls (`sec:s3-continuous`)

The continuous companion to the categorical main-text test. DV = investor trust (`Tru_diff`);
penalised smooth on the balance of sacred vs non-sacred values the trustee revealed (`NetRev`).
The SI reports the smooth as effectively linear (edf ≈ 1), smooth F = 5.42, p = .024, with a
one-SD increase in the net balance worth about +8 pp of the endowment; the association holds
controlling separately for self-similarity (p = .032) and greed (p = .027) and weakens to
marginal under both jointly (p = .069). The categorical test above is the more robust to controls.

Run on the **trustee frame** with `NetRev`, faithful to the archived
`regress Tru_diff NetRev ... if role_a_b == 2`. Model 1 is numerically identical to a smooth on
`partnerNetRev` over the investor frame --- the same dyads, with the trustee's net revealed
values stored on the trustee row (`NetRev`) rather than the investor row (`partnerNetRev`).

```{r gam-trust}
g4_m1 <- gam(Tru_diff ~ s(NetRev, k = 3),
             data = trustees, method = "REML")
g4_m2 <- gam(Tru_diff ~ s(NetRev, k = 3) + wsim1 + wsim2 + wsim3,
             data = trustees, method = "REML")
g4_m3 <- gam(Tru_diff ~ s(NetRev, k = 3) + wsim1 + wsim2 + wsim3 + Greed + age + men,
             data = trustees, method = "REML")

t4 <- bind_rows(
  data.frame(Model = "1: smooth only",            as.data.frame(gam_smooth(g4_m1))),
  data.frame(Model = "2: + self-similarity",      as.data.frame(gam_smooth(g4_m2))),
  data.frame(Model = "3: + Greed, age, gender",   as.data.frame(gam_smooth(g4_m3)))
)
kable(t4, digits = 3,
      caption = "Continuous net-balance fit (SI sec:s3-continuous) --- investor trust on the trustee's net sacred values revealed. `edf` is the smooth's effective degrees of freedom (≈1 ⇒ effectively linear). Model 1 is the smooth the SI reports (F = 5.42, p = .024); the controlled models are shown for completeness.")
```

**Percentage-point interpretation and controls.** `Tru_diff` is the change in the *share* of the
(4-unit) endowment transferred, so the linear `NetRev` slope (SI alt-specs col. 4) times 100 is
the pp figure, and the one-SD *B* below is on the same scale. The control p-values are the OLS
`NetRev` coefficient with each control set added, matching the SI's .032 / .027 / .069.

```{r maintext-trust}
b_net  <- coef(lm(Tru_diff ~ NetRev, data = trustees))["NetRev"]
sd_net <- sd(trustees$NetRev, na.rm = TRUE)
cat(sprintf("Trust: +%.2f pp of endowment per net value revealed; +%.1f pp per 1 SD (%.1f values).\n",
            100 * b_net, 100 * b_net * sd_net, sd_net))
ctrl <- function(f) { co <- summary(lm(f, data = trustees))$coefficients["NetRev", ]
  sprintf("b = %.4f, p = %.3f", co["Estimate"], co["Pr(>|t|)"]) }
cat("  + self-similarity :", ctrl(Tru_diff ~ NetRev + wsim1 + wsim2 + wsim3),          "(SI p = .032)\n")
cat("  + greed           :", ctrl(Tru_diff ~ NetRev + Greed),                          "(SI p = .027)\n")
cat("  + both jointly    :", ctrl(Tru_diff ~ NetRev + wsim1 + wsim2 + wsim3 + Greed),  "(SI p = .069)\n")
```

## SI alternative specifications (`s3-alt-specs`) --- OLS specifications of investor trust

The SI's *Alternative specifications of regression model for trustworthiness* (table
`s3-alt-specs`) reports five plain-OLS specifications of the same dependent variable --- the
change in investor trust between games (`Tru_diff`, the SI's "ΔTrust") --- on different
summaries of what the trustee revealed. These are the straight-line
companions to the continuous NetRev smooth above. The two that carry the headline signalling result at *p* < .05
are column (4), trust on the **net** balance revealed (`NetRev` = sacred − non-sacred), and
column (5), the same with `Greed` partialled out. The originally planned `SacRev`-only model is
column (1) --- the weak specification the SI's own §3.b flags and moves away from. All five are
run on the trustee frame (`role_a_b == 2`), faithful to the archived `regress Tru_diff ... if
role_a_b == 2`. Unlike the preregistered H2 port in the appendix, these are results the
manuscript stands behind, so they carry audit targets.

```{r si-table7}
# Five OLS specifications of investor trust change (Tru_diff = ΔTrust) on what the trustee
# revealed -- the linear-model companions to the continuous NetRev GAM. Cols match SI s3-alt-specs:
# (1) SacRev; (2) SacRev + total revealed (NumRev); (3) NonSacRev; (4) NetRev; (5) NetRev + Greed.
t7_specs <- list(
  "(1) SacRev"                  = Tru_diff ~ SacRev,
  "(2) SacRev + total revealed" = Tru_diff ~ SacRev + NumRev,
  "(3) NonSacRev"               = Tru_diff ~ NonSacRev,
  "(4) NetRev"                  = Tru_diff ~ NetRev,
  "(5) NetRev + Greed"          = Tru_diff ~ NetRev + Greed
)
t7_fits  <- lapply(t7_specs, function(f) lm(f, data = trustees))

# Tidy table: the focal predictor reported in each published column, with SE, t, p, R^2, N.
t7_focal <- c("SacRev", "SacRev", "NonSacRev", "NetRev", "NetRev")
t7_tab <- bind_rows(Map(function(fit, nm, foc) {
  co <- summary(fit)$coefficients
  data.frame(Model = nm, Focal = foc,
             b = co[foc, "Estimate"], SE = co[foc, "Std. Error"],
             t = co[foc, "t value"], p = co[foc, "Pr(>|t|)"],
             R2 = summary(fit)$r.squared, N = length(fit$residuals))
}, t7_fits, names(t7_fits), t7_focal))
kable(t7_tab, digits = c(0, 0, 5, 5, 2, 4, 3, 0),
      caption = "SI table `s3-alt-specs` reproduced --- OLS of investor trust change (Tru_diff) on revealed values. `Focal` is the predictor reported in the published column. Published focal coefficients: (1) .0120, (2) .0158*, (3) -.0140**, (4) .00850*, (5) .00766* (* p < .05, ** p < .01).")
```

The audit below checks the five published focal coefficients, plus the auxiliary
total-revealed coefficient in column (2), the `Greed` coefficient in column (5), and every
column's *R*², against the printed `s3-alt-specs` values.

```{r si-table7-audit}
foc_b    <- setNames(t7_tab$b, t7_tab$Model)
r2       <- setNames(t7_tab$R2, t7_tab$Model)
t7_greed <- summary(t7_fits[["(5) NetRev + Greed"]])$coefficients["Greed", "Estimate"]
t7_total <- summary(t7_fits[["(2) SacRev + total revealed"]])$coefficients["NumRev", "Estimate"]
t7_audit <- bind_rows(
  audit_row("(1) SacRev coef",         foc_b[["(1) SacRev"]],                  0.0120,  0.001, 4),
  audit_row("(2) SacRev coef",         foc_b[["(2) SacRev + total revealed"]], 0.0158,  0.001, 4),
  audit_row("(2) total-revealed coef", t7_total,                             -0.0125,  0.001, 4),
  audit_row("(3) NonSacRev coef",      foc_b[["(3) NonSacRev"]],              -0.0140,  0.001, 4),
  audit_row("(4) NetRev coef",         foc_b[["(4) NetRev"]],                  0.00850, 0.001, 5),
  audit_row("(5) NetRev coef",         foc_b[["(5) NetRev + Greed"]],          0.00766, 0.001, 5),
  audit_row("(5) Greed coef",          t7_greed,                              -0.174,   0.01,  3),
  audit_row("(1) R^2",                 r2[["(1) SacRev"]],                     0.028,   0.005, 3),
  audit_row("(2) R^2",                 r2[["(2) SacRev + total revealed"]],    0.058,   0.005, 3),
  audit_row("(3) R^2",                 r2[["(3) NonSacRev"]],                  0.057,   0.005, 3),
  audit_row("(4) R^2",                 r2[["(4) NetRev"]],                     0.054,   0.005, 3),
  audit_row("(5) R^2",                 r2[["(5) NetRev + Greed"]],             0.060,   0.005, 3)
)
kable(t7_audit, align = "lrrl")
```

**Reading the result.** All five columns reproduce to the published digits. The two `NetRev`
specifications --- (4) `NetRev` alone (*b* = .0085, *p* = .011) and (5) `NetRev` + `Greed`
(*b* = .0077, *p* = .027) --- are both significant at *p* < .05, and `Greed` is far from
significant (*p* = .39), so partialling it out leaves the signalling effect intact. The
preregistered `SacRev`-only model (1) is the weak one (*p* = .068); switching to the *balance*
revealed is what delivers significance and roughly doubles *R*² (.028 → .054). This is the
linear-model telling of the continuous NetRev GAM finding --- and since that smooth's edf ≈ 1
reported the relationship as effectively straight, the OLS line here is the same result, plainly stated.

## Trustworthiness GAM (robustness on the Figure 2 smooth)

DV = trustworthiness (`Wor_agg_avg`, proportion returned averaged across games); predictor
`SacPers` (values held), the Figure 2 measure. Nested GAMs: Model 1 (smooth only, = Figure 2),
Model 2 (+ Greed, age, gender), as a robustness check on the correlation and the +0.31 pp/value
main-text slope. Because edf ≈ 1 the smooth is effectively a straight line, so the GAM and the
linear percentage-point coefficient tell the same story.

```{r gam-worthy}
winner <- "SacPers"                       # trustworthiness predictor, by project decision
fw <- function(rhs) as.formula(sprintf("Wor_agg_avg ~ s(%s, k = 3)%s", winner, rhs))
g5_m1 <- gam(fw(""),                       data = trustees, method = "REML")
g5_m2 <- gam(fw(" + Greed + age + men"),   data = trustees, method = "REML")

t5 <- bind_rows(
  data.frame(Model = "1: smooth only (= Figure 2)", as.data.frame(gam_smooth(g5_m1))),
  data.frame(Model = "2: + Greed, age, gender",     as.data.frame(gam_smooth(g5_m2)))
)
kable(t5, digits = 3,
      caption = "Trustworthiness GAM on SacPers (values held): Model 1 is the Figure 2 smooth, Model 2 adds controls. edf ≈ 1 ⇒ effectively linear, so the linear +0.31 pp/value slope the main text reports is the same result.")

g5_greed <- summary(g5_m2)$p.table["Greed", "Estimate"]
cat(sprintf("Greed coefficient in Model 2: b = %.3f (the control does not overturn the SacPers effect).\n", g5_greed))
```

## Are the smooths non-linear? (edf + partial-effect plots)

A GAM's job is to *relax* the straight-line assumption; whether it found curvature is read off
the effective degrees of freedom (`edf`): ≈ 1 means the fitted smooth is essentially a line.
Both smooths come back at edf ≈ 1, so the linear coefficients the manuscript reports are adequate.

```{r gam-plots, fig.width=9, fig.height=4.2, fig.cap="Fitted GAM smooths with 95% confidence bands. Left: investor trust on the trustee's net sacred values revealed (continuous NetRev fit). Right: trustworthiness on sacred values held."}
par(mfrow = c(1, 2))
plot(g4_m1, shade = TRUE, shade.col = "#1a73e833", col = "#1a73e8",
     seWithMean = TRUE, main = "Continuous NetRev fit: investor trust", xlab = "NetRev")
plot(g5_m1, shade = TRUE, shade.col = "#1a73e833", col = "#1a73e8",
     seWithMean = TRUE, main = "Trustworthiness GAM", xlab = winner)
par(mfrow = c(1, 1))

cat(sprintf("Smooth edf -- NetRev fit: %.2f; trustworthiness: %.2f.\n",
            gam_smooth(g4_m1)$edf, gam_smooth(g5_m1)$edf))
cat("edf near 1 indicates the relationships are effectively linear.\n")
```

## Reported one-SD effects

For the continuous NetRev fit, the one-SD effect is *B* ≈ .078 --- the "+8 percentage points of
the endowment per one-SD increase in the net values revealed" the SI reports. Recomputed from the
fitted GAMs (predicted outcome change for a one-SD shift, covariates at their means).

```{r gam-B}
b4_m1 <- gam_oneSD(g4_m1, "NetRev")
b4_m3 <- gam_oneSD(g4_m3, "NetRev")
b5_m1 <- gam_oneSD(g5_m1, winner)
b5_m2 <- gam_oneSD(g5_m2, winner)
b_tab <- tibble(
  Quantity = c("Continuous NetRev fit one-SD B", "NetRev fit + controls one-SD B",
               "Trustworthiness M1 one-SD B (SacPers)", "Trustworthiness M2 one-SD B (SacPers)"),
  Computed = c(b4_m1, b4_m3, b5_m1, b5_m2),
  # SI reports +8 pp/SD (~.078) for the continuous NetRev fit; the others are descriptive.
  Reported = c(0.078, NA, NA, NA))
kable(b_tab, digits = 3, caption = "One-SD predicted effects recomputed from the GAMs. Reported = the manuscript's +8 pp/SD for the continuous NetRev fit; other rows descriptive.")
```

## Primary audit (main text + continuous fit)

Every load-bearing Study 3 number in the main text, plus the SI continuous fit, against the value
the manuscript prints. The remaining SI-only quantities are audited in the SI audit table below.

```{r audit}
g4_1 <- gam_smooth(g4_m1)
s3_audit <- bind_rows(
  # Warranted trustworthiness (Figure 2)
  audit_row("Trustworthiness: pp per value held (main text +0.31)", 100 * b_sp,       0.31,  0.03, 2),
  audit_row("Trustworthiness: pp per 1 SD (main text +5.7)",        100 * b_sp * sd_sp, 5.7, 0.3,  1),
  audit_row("Trustworthiness: Pearson r (descriptive)",            s3_return$r,       NA,    0.03, 3),
  # Investor categorical test (primary)
  audit_row("Binary investor: b (SI 0.196; main text ~20 pp)",     bin_b,             0.196, 0.01, 3),
  audit_row("Binary investor: CLES (main text .65)",               bin_cles,          0.65,  0.02, 2),
  audit_row("Binary investor: R^2 (main text ~8%)",                bin_r2,            0.08,  0.02, 3),
  audit_row("Binary investor + controls: b (main text p=.01)",     binc_b,            0.167, 0.01, 3),
  audit_row("Binary investor + NonSacRev: b (SI 0.15)",            binn_b,            0.15,  0.01, 3),
  audit_row("Binary investor + NonSacRev + NumRev: b (SI 0.20)",   binnn_b,           0.20,  0.01, 3),
  audit_row("Binary investor on belief 'trusts more': b (SI 0.97)",binbel_b,          0.97,  0.03, 3),
  # Trustee side is graded (binary null) -- exploration values, descriptive
  audit_row("Trustee binary held>0: b in pp (exploration +6.4)",   100 * tb_b,        NA,    0.5,  1),
  audit_row("Trustee binary: CLES (exploration .59)",              tb_cles,           NA,    0.02, 3),
  # Median split recovers the effect (main text + SI sec:s3-binary)
  audit_row("Trustee median split: b in pp (SI +10.9)",            100 * mb_b,        10.9,  0.5,  1),
  audit_row("Trustee median split: p (SI < .001)",                 mb_p,              NA,    0.001, 4),
  audit_row("Trustee median split: CLES (SI .68)",                 mb_cles,           0.68,  0.02, 2),
  # Continuous fit (SI sec:s3-continuous)
  audit_row("Continuous NetRev smooth F (SI 5.42)",                g4_1$F,            5.42,  0.1),
  audit_row("Continuous NetRev one-SD B (SI +8 pp/SD)",            b4_m1,             0.078, 0.02, 3)
)
kable(s3_audit, align = "lrrl")
```

**Reading the result.** The substantive findings reproduce. Warranted trustworthiness (Figure 2):
trustees who *hold* more sacred values return more, +0.31 pp per value held (+5.7 pp/SD),
edf ≈ 1. The primary investor result is categorical: crossing the sacred-revelation threshold
raises trust by ≈ 20 pp (b = .196, CLES = .65, ~8% of variance), robust to controls (b = .167,
p = .01) and to holding the non-sacred count fixed. The trustee-side binary is null --- the
warranted side is genuinely graded --- and the continuous NetRev fit (SI) reproduces at F = 5.42,
one-SD B ≈ .078. Trustee-side binary rows carry no printed manuscript target (the paper makes only
the qualitative null claim), so they are marked descriptive and checked against the exploration.

> **Note --- two predictors by design.** The investor signal is read categorically off the
> *balance revealed* (`SacRev > 0`), the noisy partial disclosure a receiver actually sees;
> trustworthiness is modelled on sacred values *held* (`SacPers`), the complete trait. Different
> operationalisations by design, not an inconsistency --- the reception-vs-trait asymmetry.

> **Note --- controls are the trustee's.** The archived `.do` runs the controlled models
> `... if role_a_b == 2`, so Greed/age/gender are the trustee's.

---

## SI-only analyses

> **Scope.** The subsections below reproduce every analysis the paper reports **only in the
> Supporting Information** for Study 3: the post-game belief correlations, the
> willingness-to-revise correlation, the revelation-preference analysis (preregistered
> chi-square and the mixed-effects logit), the trust-connoting content bake-off and its
> factor-loading companion, and the religiosity rank-sum table. Each computed value is
> collected into the SI audit table at the end of the section.

```{r si-load}
# Extra archives used only by the SI analyses below.
tc    <- read_dta(file.path(S3_DIR, "trust-combined.dta"))            # 242, trust game
means <- read_dta(file.path(S3_DIR, "combined-ult-trust-means.dta"))  # per-statement endorsement rates
```

## Post-game beliefs about trustworthiness and trust (`s3-scatter-belief`, `s3-scatter-trustworthiness`)

After the second trust game but before learning either outcome, participants reported whether
they expected their second-game partner to be more or less trusting / trustworthy than their
first-game partner. The SI correlates these beliefs against the balance of sacred versus
non-sacred values revealed and finds the same pattern as the incentivized behaviour:

- **Belief the partner will trust more** (`blf_trustee_trust`, held by the trustee who did the
  revealing) against the trustee's own net values revealed (`NetRev`): the strong correlation
  in `s3-scatter-belief`.
- **Belief the partner will be more trustworthy** (`blf_investor_worthy`, held by the investor)
  against the balance the trustee partner revealed (`partnerNetRev`): the weaker,
  marginal correlation in `s3-scatter-trustworthiness`.

```{r si-beliefs}
belief_trust  <- pearson_report(as.numeric(trustees$blf_trustee_trust),  trustees$NetRev)
belief_worthy <- pearson_report(as.numeric(investors$blf_investor_worthy), investors$partnerNetRev)

cat(sprintf("s3-scatter-belief          belief 'partner trusts more' ~ own NetRev     : r = %.3f, p = %.4f (n = %d)\n",
            belief_trust$r,  belief_trust$p,  belief_trust$n))
cat(sprintf("s3-scatter-trustworthiness belief 'partner worthier'  ~ partner NetRev  : r = %.3f, p = %.4f (n = %d)\n",
            belief_worthy$r, belief_worthy$p, belief_worthy$n))
```

## Willingness to revise answers (`s3-survey-revision`)

The SI reports that participants holding more sacred values are less inclined to revise their
inventory answers. The archived item `survey1playerexp_ques1` ("would give identical answers in
future") runs in the *opposite* direction to "willingness to revise", so a **positive**
correlation with sacred values held (`SacPers`) is the SI's finding: more sacred values, more
likely to stand pat. Defined for all trust-game participants (one is missing the item, so
n = 241).

```{r si-revision}
revision <- pearson_report(as.numeric(full$survey1playerexp_ques1), full$SacPers)
cat(sprintf("s3-survey-revision  'would give identical answers' ~ sacred values held : r = %.3f, p = %.5f (n = %d)\n",
            revision$r, revision$p, revision$n))
```

## The revelation procedure (`sec:s3-revelation-procedure`)

The SI's revelation-procedure subsection states that a random draw from {0, 2, 5, 9, 16}
capped how many of the eighteen rankable
statements could be concealed, **so every trustee revealed at least two statements**. That
factual claim is checkable against the archived per-trustee revealed count (`NumRev`).

```{r si-revelation-procedure}
concealed <- 18 - trustees$NumRev
cat("Concealed statements per trustee (18 - NumRev):\n"); print(table(concealed))
cat(sprintf("Revealed range: %d--%d (n = %d trustees)\n",
            min(trustees$NumRev), max(trustees$NumRev), nrow(trustees)))
stopifnot("SI claim 'every trustee revealed at least two statements' fails" =
            min(trustees$NumRev, na.rm = TRUE) >= 2,
          "Concealed count exceeds the maximum draw of 16" =
            max(concealed, na.rm = TRUE) <= 16)
# Note: concealed counts need not sit exactly on {0, 2, 5, 9, 16} -- the draw caps how many
# of a trustee's ranked statements are withheld, and trustees could rank fewer than the cap.
```

## Did trustees reveal sacred values preferentially? (`s3-mixed-logit`)

The paper's third, tentative hypothesis was that trustees would preferentially reveal their
*sacred* values. The originally preregistered test --- a simple chi-square on the pooled
statement-level reveal decisions --- was significant (58% of sacred values revealed versus 45%
of non-sacred; the SI reports chi-square(1) = 28.9). But that test ignores the non-independence
of repeated statements and repeated participants. The SI therefore refits the reveal decision
with a **mixed-effects logistic regression**, crossed random intercepts for participant and for
statement (`s3-mixed-logit`). Under that conservative model the sacredness coefficient is small,
*negative* (if anything, sacred statements are concealed slightly more), and not significant.

```{r si-reveal-build}
# Reshape to one row per (trustee x rankable statement); flag whether each was revealed and
# whether it was held sacred. Ports `hypothesis 3 code.do`. 121 trustees x 18 statements = 2178.
rankable <- c(6,8,10,12,18,20,24,25,27,32,35,37,42,48,54,56,58,60)
num <- function(v) suppressWarnings(as.numeric(v))
rows <- which(tc$role_a_b == 2)
reveal_long <- do.call(rbind, lapply(rows, function(i) {
  rank_x  <- num(unlist(tc[i, paste0("t3_conceal_rank_", 1:18)]))
  price_x <- num(unlist(tc[i, paste0("t3_raw_price_", 1:18)]))
  abs_x   <- num(unlist(tc[i, paste0("t3_conceal", 1:18, "_abs")]))
  endorse <- num(unlist(tc[i, paste0("t2_a_sp", rankable)])) - 1     # 1/2 -> 0/1
  sp_rank <- vapply(rankable, function(y) { p <- which(rank_x == y + 1); if (length(p)) p[1] else NA_integer_ }, integer(1))
  sp_sac  <- vapply(rankable, function(y) as.integer(any(price_x == 101 & abs_x == y, na.rm = TRUE)), integer(1))
  data.frame(pid = i, sp_num = rankable, sp_rank = sp_rank, sp_sacred = sp_sac, endorse = endorse)
}))
reveal_long$revealed <- as.integer(is.na(reveal_long$sp_rank))       # not ranked => revealed
reveal_long <- merge(reveal_long, means[, c("sp_num", "avg_endorse")], by = "sp_num")
reveal_long$deviance <- (reveal_long$endorse - reveal_long$avg_endorse)^2
reveal_long$pid <- factor(reveal_long$pid); reveal_long$spf <- factor(reveal_long$sp_num)
cat(sprintf("Statement-level rows: %d  (%.0f%% of sacred revealed, %.0f%% of non-sacred revealed)\n",
            nrow(reveal_long),
            100 * mean(reveal_long$revealed[reveal_long$sp_sacred == 1]),
            100 * mean(reveal_long$revealed[reveal_long$sp_sacred == 0])))
```

```{r si-reveal-chi}
# Preregistered chi-square, WITHOUT continuity correction -- matches the SI's 28.9 (and Stata's
# `tabulate, chi2`, which does not apply Yates's correction). The corrected value is ~28.3.
reveal_tab <- table(sacred = reveal_long$sp_sacred, revealed = reveal_long$revealed)
reveal_chi <- suppressWarnings(chisq.test(reveal_tab, correct = FALSE))
cat(sprintf("Preregistered chi-square (uncorrected): X2(%d) = %.2f, p = %.4f\n",
            reveal_chi$parameter, reveal_chi$statistic, reveal_chi$p.value))
```

```{r si-reveal-logit}
m_rev1 <- glmer(revealed ~ sp_sacred + (1 | pid) + (1 | spf), data = reveal_long,
                family = binomial, control = glmerControl(optimizer = "bobyqa"))
m_rev2 <- glmer(revealed ~ sp_sacred + deviance + (1 | pid) + (1 | spf), data = reveal_long,
                family = binomial, control = glmerControl(optimizer = "bobyqa"))
mlogit_tab <- rbind(
  data.frame(Model = "(1) revealed ~ sacred",            as.data.frame(summary(m_rev1)$coefficients)["sp_sacred", ]),
  data.frame(Model = "(2) revealed ~ sacred + deviance", as.data.frame(summary(m_rev2)$coefficients)["sp_sacred", ]))
kable(mlogit_tab, digits = 3, row.names = FALSE,
      caption = "SI table `s3-mixed-logit` reproduced --- effect of a statement being sacred on the log-odds of revealing it. Both models carry crossed random intercepts for participant and statement.")
rev1_b <- summary(m_rev1)$coefficients["sp_sacred", "Estimate"]
rev1_p <- summary(m_rev1)$coefficients["sp_sacred", "Pr(>|z|)"]
rev2_b <- summary(m_rev2)$coefficients["sp_sacred", "Estimate"]
cat(sprintf("Model 1 sacred coef b = %.3f (p = %.3f); Model 2 sacred coef b = %.3f.\n",
            rev1_b, rev1_p, rev2_b))
```

> **glmer vs Stata `melogit`.** The fixed-effect coefficients reproduce the SI to three
> decimals (b = -0.269 / -0.245), as do the variance components (participant ≈ 21, statement
> ≈ 0.32). The *p*-value the SI quotes for sacredness (p = 0.264) comes from Stata's
> `melogit` Wald test; `lme4::glmer` returns a slightly different *z*/*p* for the same
> coefficient. The substantive reading --- a small, non-significant, wrong-signed effect ---
> is identical. The coefficient (not the *p*) is what the audit targets.

## Did specific trust-implying statements explain the result? (SI bake-off)

A natural alternative to the sacred-values account is that the Study 3 effect was driven not
by sacred values in general but by the *content* of a handful of inventory statements that, on
their face, imply the speaker is trustworthy. The OSF `hypothesis 2 code.do` (its "trust
implying value statements" block) builds four indices from the seven trust-connoting
statements and regresses the outcomes on each one *singly*; the SI then compares them, loosely,
to the general sacred-values relationship. This section sharpens that comparison into an
explicit head-to-head and reproduces the SI table.

The four indices, exactly as constructed in `import data.do` (lines 409--426):

- `trust_sp_index` --- number of the seven trust-connoting statements the trustee **endorsed**
  (0--7). Defined for all 121 trustees.
- `trust_sp_sacred_index` --- of those, the number also **held sacred** (0--7).
- `trust_revealed_index` --- of the **two** trust-connoting statements eligible to be revealed
  before the second trust game, the number the trustee **revealed** in the trustworthy
  direction (0--2).
- `trust_sacred_revealed_index` --- of those two, the number revealed *and* held sacred (0--2).

**The n = 88 quirk.** The two revealed indices are built as a sum of two per-statement terms,
each left missing when its statement was concealed. In Stata (and so in the archived `.dta`)
any arithmetic touching a missing value returns missing, so the index goes missing whenever
*either* eligible statement was concealed --- it is defined only for trustees who revealed
**both**. That is 88 of 121 (91 revealed statement 25, 100 revealed statement 33, 88 revealed
both; "at least one" would have been 103). The revealed-content analyses therefore rest on
*n* = 88, against *n* = 121 for the held-content and benchmark measures. The check below
confirms the count.

```{r si-trust-ncheck}
r25 <- !is.na(trustees$trust_sp_25_rev)
r33 <- !is.na(trustees$trust_sp_33_rev)
cat(sprintf("revealed stmt 25: %d | stmt 33: %d | BOTH: %d | AT LEAST ONE: %d\n",
            sum(r25), sum(r33), sum(r25 & r33), sum(r25 | r33)))
cat(sprintf("non-missing trust_revealed_index: %d  (= revealed both)\n",
            sum(!is.na(trustees$trust_revealed_index))))
```

### Helpers

Predictors are on different scales (0--7, 0--2, counts), so the like-for-like comparison is the
**standardized** coefficient β; in a single-predictor model *R*² is just β². `uni_fit` returns
the standardized simple regression for one predictor; `headtohead` adds each content index to
the outcome's general sacred-values benchmark (sacred values held for trustworthiness, net
sacred values revealed for change in trust) and reports the index's coefficient before and
after that adjustment. Every model uses the rows complete on the variables it involves, so the
*alone* and *adjusted* coefficients for a given index share one sample.

```{r si-trust-helpers}
# Standardized simple regression of dv on a single predictor xv. ----------------
uni_fit <- function(data, dv, xv) {
  d <- data[stats::complete.cases(data[[dv]], data[[xv]]), ]
  m <- lm(scale(d[[dv]])[, 1] ~ scale(d[[xv]])[, 1])
  co <- summary(m)$coefficients
  tibble(Predictor = xv, n = nrow(d), beta = co[2, "Estimate"],
         p = co[2, "Pr(>|t|)"], R2 = summary(m)$r.squared)
}

# Univariate bake-off over several predictors, sorted by |beta|. ---------------
bakeoff <- function(data, dv, predictors) {
  tab <- bind_rows(lapply(predictors, function(p) uni_fit(data, dv, p)))
  tab[order(-abs(tab$beta)), ]
}

# Head-to-head table (reproduces SI Table s3-trust-index): each index alone, then
# its standardized coefficient once the benchmark is added to the model. --------
headtohead <- function(data, dv, benchmark, indices) {
  std <- function(v) scale(v)[, 1]
  # benchmark row: univariate only
  bench <- uni_fit(data, dv, benchmark)
  rows <- list(tibble(Predictor = benchmark, n = bench$n, beta = bench$beta,
                      p = bench$p, beta_adj = NA_real_, p_adj = NA_real_))
  for (xv in indices) {
    u <- uni_fit(data, dv, xv)
    d <- data[stats::complete.cases(data[[dv]], data[[xv]], data[[benchmark]]), ]
    m <- lm(std(d[[dv]]) ~ std(d[[benchmark]]) + std(d[[xv]]))
    co <- summary(m)$coefficients
    rows[[length(rows) + 1]] <- tibble(
      Predictor = xv, n = u$n, beta = u$beta, p = u$p,
      beta_adj = co[3, "Estimate"], p_adj = co[3, "Pr(>|t|)"])
  }
  bind_rows(rows)
}
```

### Trustworthiness (`Wor_agg_avg`)

```{r si-trust-worthy}
worthy_uni <- bakeoff(trustees, "Wor_agg_avg",
                      c("SacPers", "NetRev", "trust_sp_index", "trust_sp_sacred_index"))
kable(worthy_uni, digits = 3,
      caption = "Trustworthiness: univariate standardized bake-off (sorted by |beta|). R2 = beta^2.")

worthy_h2h <- headtohead(trustees, "Wor_agg_avg", "SacPers",
                         c("trust_sp_index", "trust_sp_sacred_index"))
kable(worthy_h2h, digits = 3,
      col.names = c("Predictor", "n", "beta", "p", "beta (adj.)", "p (adj.)"),
      caption = "Trustworthiness head-to-head. `adj.` = the index's standardized coefficient with sacred values held (SacPers) in the model. Reproduces the trustworthiness panel of SI Table s3-trust-index.")
```

### Change in trust (`Tru_diff`)

```{r si-trust-trust}
trust_uni <- bakeoff(trustees, "Tru_diff",
                     c("NetRev", "trust_revealed_index", "trust_sacred_revealed_index"))
kable(trust_uni, digits = 3,
      caption = "Change in trust: univariate standardized bake-off (sorted by |beta|). Revealed indices defined on n = 88.")

trust_h2h <- headtohead(trustees, "Tru_diff", "NetRev",
                        c("trust_revealed_index", "trust_sacred_revealed_index"))
kable(trust_h2h, digits = 3,
      col.names = c("Predictor", "n", "beta", "p", "beta (adj.)", "p (adj.)"),
      caption = "Change-in-trust head-to-head. `adj.` = the index's standardized coefficient with net sacred values revealed (NetRev) in the model. Reproduces the change-in-trust panel of SI Table s3-trust-index.")
```

**Reading the result.** The general sacred-values measure wins on both outcomes. For
trustworthiness, sacred values held is the strongest single predictor (β ≈ .34) and neither
content index adds anything once it is included: the held trust index is null to begin with,
and the sacred trust index --- respectable alone (β ≈ .27) --- collapses to roughly zero with
`SacPers` in the model. For change in trust, the net balance revealed is the only measure
significant on the full sample (β ≈ .23, *p* ≈ .01); the revealed content indices are weaker,
rest on the smaller *n* = 88, and lose their significance once the net balance is partialled
out. In no case does specific trust-connoting content explain variance the general
sacred-values signal does not. These are the numbers behind SI Table `s3-trust-index`.

### Factor loadings versus the aggregate

The SI also checks whether any empirically identified *factor* of the inventory (Section
`s3-factors`) predicts trustworthiness better than the raw aggregate. It reports that a high
loading on the first factor goes with greater trustworthiness, but explains **less** variance
than the total sacred-values count (9.2% versus 11.8%), and that no revealed factor loading
predicts change in trust. Here `full_f1` is the archived first-factor score (full inventory)
and `SacPers` the aggregate.

```{r si-factor-r2}
r2_f1  <- summary(lm(Wor_agg_avg ~ full_f1, data = trustees))$r.squared
r2_agg <- summary(lm(Wor_agg_avg ~ SacPers, data = trustees))$r.squared
r2_rf1 <- summary(lm(Tru_diff    ~ rev_f1,  data = trustees))$r.squared
cat(sprintf("Trustworthiness R^2: first factor (full_f1) = %.1f%% vs aggregate (SacPers) = %.1f%%\n",
            100 * r2_f1, 100 * r2_agg))
cat(sprintf("Change in trust R^2: first revealed factor (rev_f1) = %.1f%% (SI: no revealed factor predicts it)\n",
            100 * r2_rf1))
```

Both figures reproduce exactly: the aggregate gives 11.8% and the archived first-factor
score (`full_f1`) gives 9.24% -> 9.2%. The factor explains less than the aggregate.

## Religiosity, trust and trustworthiness (`s3-religiosity`)

The SI reports that religiosity --- under three measures (any religious affiliation, belief in
God, and treating religion itself as sacred) --- is unrelated to trust or trustworthiness. The
single significant relationship is that people who hold *religion* sacred hold many more sacred
values overall (Mann--Whitney *z* = 6.895, *p* < .0001). The table below reproduces the group
means (`s3-religiosity`); the *z*-test is on the number of sacred values by the
religion-is-sacred split, over all 242 trust-game participants.

```{r si-religiosity}
relig_desc <- function(gvar) {
  g <- droplevels(as_factor(tc[[gvar]]))
  data.frame(
    Measure   = gvar,
    Group     = levels(g),
    `Sacred values`      = round(tapply(tc$SacPers,     g, mean, na.rm = TRUE), 1),
    `Trustworthy (g1)`   = round(tapply(tc$Wor_pre_avg, g, mean, na.rm = TRUE), 2),
    `Trust received`     = round(tapply(tc$Tru_post,    g, mean, na.rm = TRUE), 2),
    N         = as.integer(tapply(tc$SacPers, g, function(z) sum(!is.na(z)))),
    check.names = FALSE, row.names = NULL)
}
relig_tab <- bind_rows(lapply(c("religious_bin", "believe_god", "religious_sacred"), relig_desc))
kable(relig_tab, digits = 2,
      caption = "SI table `s3-religiosity` reproduced --- sacred values, trustworthiness (game 1 proportion returned), and trust received, by religious affiliation / belief in God / religion-is-sacred.")

# The one tested effect: sacred values by the religion-is-sacred split.
rs <- factor(as_factor(tc$religious_sacred))
relig_z <- as.numeric(coin::statistic(
  coin::wilcox_test(tc$SacPers ~ rs, distribution = "asymptotic"), "standardized"))
cat(sprintf("Sacred values ~ religion-is-sacred: Mann-Whitney |z| = %.3f, means %.1f (sacred) vs %.1f (not)\n",
            abs(relig_z),
            mean(tc$SacPers[tc$religious_sacred == 1], na.rm = TRUE),
            mean(tc$SacPers[tc$religious_sacred == 0], na.rm = TRUE)))
```

> **Descriptive flag.** The sacred-values row and the *N* column reproduce the SI table
> exactly. The trustworthiness-in-game-1 and trust-received rows reproduce the SI's pattern but
> a few cells differ by ~.02--.05 from the printed figures (e.g. belief-in-God trustworthiness
> .36/.33 here vs .39/.31 in the SI). These are untested descriptive means; the discrepancy is
> logged for review, not silently reconciled.

## SI audit: computed vs SI

Every value reported in the Study 3 SI, collected against its printed figure.

```{r si-audit}
# Pull a standardized coefficient (alone or adjusted) out of a head-to-head table by predictor.
grab <- function(tab, pred, col = "beta") tab[[col]][tab$Predictor == pred]

si_audit <- bind_rows(
  # Beliefs
  audit_row("Belief 'partner trusts more' r (s3-scatter-belief)",       belief_trust$r,  0.428, 0.02, 3),
  audit_row("Belief 'partner worthier' r (s3-scatter-trustworthiness)", belief_worthy$r, 0.176, 0.02, 3),
  audit_row("Belief 'partner worthier' p",                              belief_worthy$p, 0.054, 0.01, 3),
  # Revision
  audit_row("Revise-answers r (s3-survey-revision)",                    revision$r,      0.335, 0.02, 3),
  # Revelation preference
  audit_row("Revelation chi-square (s3-mixed-logit)",                   reveal_chi$statistic, 28.9, 0.5, 1),
  audit_row("Mixed-logit sacred coef, Model 1",                         rev1_b,         -0.269, 0.02, 3),
  audit_row("Mixed-logit sacred coef, Model 2",                         rev2_b,         -0.245, 0.02, 3),
  # Trust-connoting content (s3-trust-index)
  audit_row("Trustworthiness: SacPers beta",           grab(worthy_h2h, "SacPers"),               0.34, 0.03, 2),
  audit_row("Trustworthiness: trust index beta",       grab(worthy_h2h, "trust_sp_index"),        0.13, 0.03, 2),
  audit_row("Trustworthiness: trust-sacred index beta",grab(worthy_h2h, "trust_sp_sacred_index"), 0.27, 0.03, 2),
  audit_row("Trustworthiness: trust-sacred index beta (adj.)",
            grab(worthy_h2h, "trust_sp_sacred_index", "beta_adj"),                                -0.18, 0.04, 2),
  audit_row("Change in trust: NetRev beta",            grab(trust_h2h, "NetRev"),                  0.23, 0.03, 2),
  audit_row("Change in trust: revealed index beta",    grab(trust_h2h, "trust_revealed_index"),    0.19, 0.03, 2),
  audit_row("Change in trust: sacred-revealed index beta",
            grab(trust_h2h, "trust_sacred_revealed_index"),                                        0.22, 0.03, 2),
  audit_row("Change in trust: sacred-revealed index beta (adj.)",
            grab(trust_h2h, "trust_sacred_revealed_index", "beta_adj"),                            0.11, 0.04, 2),
  # Factor loadings vs aggregate
  audit_row("Trustworthiness R^2: first factor",       r2_f1,  0.092, 0.005, 3),
  audit_row("Trustworthiness R^2: aggregate",          r2_agg, 0.118, 0.01, 3),
  # Religiosity
  audit_row("Religiosity Mann-Whitney |z|",            abs(relig_z), 6.895, 0.05, 3)
)
kable(si_audit, align = "lrrl")
```

---

# Audit summary

```{r audit-summary, echo=FALSE}
# All three audit tables share the same columns (Quantity / Reported / Computed / Check); the
# "Reported" column holds the value the manuscript prints (main text or SI).
all_audit <- dplyr::bind_rows(
  dplyr::mutate(s3_audit, Section = "Main / continuous", .before = 1),
  dplyr::mutate(t7_audit, Section = "SI alt-specs",      .before = 1),
  dplyr::mutate(si_audit, Section = "SI",                .before = 1)
)
n_match  <- sum(all_audit$Check == "match")
n_close  <- sum(all_audit$Check == "close")
n_differ <- sum(all_audit$Check == "DIFFERS")
cat(sprintf("Audited claims: %d | match: %d | close (bootstrap/rounding/rebuild noise): %d | DIFFERS: %d\n",
            nrow(all_audit), n_match, n_close, n_differ))
if (n_differ > 0) {
  cat("\nClaims that differ from the paper:\n")
  print(knitr::kable(subset(all_audit, Check == "DIFFERS",
                            select = c("Section", "Quantity", "Reported", "Computed"))))
}
```

---

# Preregistered specifications (OSF Stata port)

> **Status: preregistered provenance.** This section ports the *preregistered* Stata
> specifications from the OSF archive --- the H1 returns-on-sacred-values OLS
> (`hypothesis 1 code.do`) and the H2 similarity-controlled OLS (`hypothesis 2 code.do`). The
> paper departs from the H2 investor specification (it reports the categorical test and, in the
> SI, the net-balance and preregistered-count models instead), so these mostly document what
> was preregistered. One exception: the SI design history and the main text cite the H2
> port's SacRev *p*-value (*p* = .2154, rounded to .22 in the main text) alongside the
> unadjusted count-model *p* = .068, so that quantity carries an audit row below. The other two preregistered tests --- the
> religiosity rank-sums and the H3 mixed logit --- *are* reported in the SI and appear as audited
> results in the SI section above, not here. Included for transparency / reproducibility of the
> OSF code.

```{r appendix-load}
tc_tru <- tc %>% filter(role_a_b == 2)   # `tc` was loaded in the SI section above
```

## H1 --- do sacred valuers return more? (OLS)

`regress Wor_agg_avg SacPers if role_a_b == 2` and variants.

```{r h1-ols}
h1 <- list(
  "returns ~ SV (aggregate)"   = lm(Wor_agg_avg ~ SacPers, data = tc_tru),
  "returns ~ SV (Game 1 only)" = lm(Wor_pre_avg ~ SacPers, data = tc_tru),
  "returns ~ SV + Greed"       = lm(Wor_agg_avg ~ SacPers + Greed, data = tc_tru)
)
h1_tab <- do.call(rbind, lapply(names(h1), function(nm) {
  co <- summary(h1[[nm]])$coefficients
  data.frame(Model = nm, b_SacPers = co["SacPers","Estimate"],
             SE = co["SacPers","Std. Error"], t = co["SacPers","t value"],
             p = co["SacPers","Pr(>|t|)"], R2 = summary(h1[[nm]])$r.squared)
}))
kable(h1_tab, digits = 3, caption = "H1: trustee returns regressed on sacred-value endorsement.")
```

> **The religiosity rank-sum tests are now an audited SI result** --- see *Religiosity, trust
> and trustworthiness* (`s3-religiosity`) above, which reproduces the SI table and the one
> significant test (Mann--Whitney *z* = 6.895). Not duplicated here.

## H2 --- preregistered test (OLS with revealed similarity)

Ports the preregistered test `regress Tru_diff SacRev Greed wtd_rev_sim_factor*; test SacRev`.
The revealed-similarity factors (`wsim1/2/3`) were constructed once in the prep section above
(min-max normalised factor scores, NumRev-weighted, dyad-matched) and are reused here, so the
OLS port and the controlled GAMs share a single construction.

```{r h2}
h2 <- lm(Tru_diff ~ SacRev + Greed + wsim1 + wsim2 + wsim3, data = trustees)
kable(as.data.frame(summary(h2)$coefficients), digits = 3,
      caption = "H2 (preregistered): change in trust on net sacred revealed, controlling for greed and revealed similarity.")
cat(sprintf("Preregistered test of SacRev: b = %.3f, t = %.2f, p = %.4f\n",
            summary(h2)$coefficients["SacRev", "Estimate"],
            summary(h2)$coefficients["SacRev", "t value"],
            summary(h2)$coefficients["SacRev", "Pr(>|t|)"]))
# This p is cited in the SI design history (p = .2154) and, rounded to .22, in the
# main text, so it is audited here.
h2_p_audit <- audit_row("H2 prereg SacRev p (SI .2154; main text .22)",
                        summary(h2)$coefficients["SacRev", "Pr(>|t|)"], 0.2154, 0.0005, 4)
kable(h2_p_audit, align = "lrrl")
stopifnot("H2 preregistered SacRev p does not reproduce" = all(h2_p_audit$Check == "match"))
```

> **The mixed-logit revelation test (former H3) is now an audited SI result** --- see
> *Did trustees reveal sacred values preferentially?* (`s3-mixed-logit`) above --- because the
> paper reports it in the SI. It is no longer duplicated here.

---

# Reproducibility record

```{r session, echo=FALSE}
si <- sessionInfo()
cat("R version:", si$R.version$version.string, "\n")
cat("Platform :", si$platform, "\n")
cat("Seed     : 20260619\n\nKey package versions:\n")
for (p in c("haven", "dplyr", "mgcv", "ggplot2", "tibble", "knitr", "rmarkdown")) {
  v <- tryCatch(as.character(packageVersion(p)), error = function(e) "not installed")
  cat(sprintf("  %-10s %s\n", p, v))
}
```

```{r session-full, echo=FALSE}
sessionInfo()
```
