---
title: "Sacred values as index signals of trustworthy character --- Reproducibility Pipeline (Studies 1 & 2)"
subtitle: "Push-button reproduction of every statistical test reported for Studies 1 and 2, in both the main text and the Supporting Information"
author: "Sacred Values project"
date: "Knitted `r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    number_sections: true
    code_folding: show
    df_print: kable
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 7,
  fig.height = 4.2
)
```

# What this document is

This is a **single, self-contained reproducibility pipeline** for Studies 1 and 2 of
*Sacred values as index signals of trustworthy character*. Knitting it from a clean R
session reproduces, from the raw OSF data, **every statistical test reported for these
two studies --- in both the main text and the Supporting Information (SI)** --- and audits
each computed value against the number printed in the paper. Concretely it covers:

- the headline trustworthiness contrast, its effect size and probability of superiority
  (CLES), and the out-group replication (main text);
- the low-sanctity-observer robustness check (main text);
- the pre-registered *parametric* tests, trait by trait and relationship by relationship (SI);
- the sensitivity analysis *without exclusions* (SI);
- the exploratory *selfishness* attribution analyses, including the single and dual
  mediation models (SI);
- the Shapiro-Wilk normality tests that motivate the non-parametric approach (SI); and
- the inter-correlation matrices among the key variables (SI), plus the Study 2
  sacralizing-of-friendship and closeness-to-community effects (SI).

**Design standard: push-button replication** (Wood, Müller & Brown, *PLOS ONE* 2018):
a third party should be able to take the archived data and code and *"push the button"*
to reproduce the published results **without contacting the authors**. Concretely, this
file is built to satisfy their criteria:

| Push-button criterion | How this document meets it |
|---|---|
| Ready-to-run, no author intervention | One `.Rmd`; `Knit` runs everything top to bottom. |
| Raw data + analysis code together | Reads the original `.sav` files directly; no hidden intermediate datasets. |
| No machine-specific paths | Data location is auto-detected by walking up the directory tree; no absolute paths, no `setwd()`. |
| Data structure matches code | Sample sizes are **asserted** (`stopifnot`) against the paper's reported *N*; the knit fails loudly if the data drifted. |
| Key results declared in advance | Reported values are written into a target table *before* the data are analysed, then compared automatically. |
| Documented software environment | Full `sessionInfo()` and package versions are printed at the end. |
| Reproducible randomness | A single global seed is set; every bootstrap additionally fixes its own seed. |

No original SPSS syntax survived, so the analyses here are a from-scratch rebuild guided
by the OSF variable keys. The fact that the rebuilt numbers match the paper (see the
audit tables) is itself the evidence that the rebuild is faithful.

---

# Setup

## Packages, seed, and data location

```{r packages}
suppressPackageStartupMessages({
  library(haven)    # read SPSS .sav
  library(dplyr)    # data manipulation
  library(coin)     # Mann-Whitney with exact standardized Z
  library(ggplot2)  # figures
  library(tibble)   # tidy result tables
  library(knitr)    # kable tables
})

# One seed for the whole document (YYYYMMDD). Every bootstrap below also fixes its own
# seed so the confidence intervals are byte-for-byte reproducible.
set.seed(20260619)
```

The data live in an `osf/` folder (a copy of the OSF archive at
<https://osf.io/ju9rq/>). Rather than hard-code a path, we locate that folder by walking
up from wherever this file is knitted. To run against data stored elsewhere, set
`DATA_DIR` by hand in the chunk below.

```{r data-root}
# Walk up the directory tree until we find a folder containing `marker`.
find_data_root <- function(start = getwd(), marker = "osf") {
  d <- normalizePath(start, winslash = "/", mustWork = FALSE)
  repeat {
    if (dir.exists(file.path(d, marker))) return(d)
    parent <- dirname(d)
    if (identical(parent, d)) {
      stop("Could not locate the '", marker, "' data directory by walking up from ",
           start, ".\n  Set DATA_DIR manually in this chunk.", call. = FALSE)
    }
    d <- parent
  }
}

DATA_DIR <- file.path(find_data_root(), "osf")   # <- override here if needed

s1_file <- file.path(DATA_DIR, "Study 1 - Sacred Ring",    "OSF SV Trust - Study 1.sav")
s2_file <- file.path(DATA_DIR, "Study 2 - Pork Investment", "OSF SV Trust - Study 2.sav")

# Fail early and clearly if the data are not where we expect.
stopifnot(
  "Study 1 .sav not found" = file.exists(s1_file),
  "Study 2 .sav not found" = file.exists(s2_file)
)
cat("Data directory:\n  ", DATA_DIR, "\n")
```

## Statistical helper functions {.tabset}

These functions do all the statistical work. They are defined here, in the open, so a
sceptical reader can audit every line. (Click the **Code** toggle to collapse them.)

```{r helpers}
# Mean of a set of items, coercing SPSS labelled columns to plain numeric -----
composite <- function(df, items) {
  rowMeans(sapply(items, function(v) as.numeric(df[[v]])), na.rm = TRUE)
}

# Cronbach's alpha (raw) for a data frame / matrix of items -------------------
cronbach_alpha <- function(df, items) {
  m <- sapply(items, function(v) as.numeric(df[[v]]))
  m <- m[stats::complete.cases(m), , drop = FALSE]
  k <- ncol(m)
  (k / (k - 1)) * (1 - sum(apply(m, 2, stats::var)) / stats::var(rowSums(m)))
}

# Mann-Whitney U test with the exact standardized Z (via coin) ----------------
# Returns both U conventions. The main text reports U_max (the larger U); a few SI
# effects are printed as U_min, so both are available. Effect size r = |Z| / sqrt(N).
mann_whitney <- function(dv, group) {
  group <- droplevels(as.factor(group))
  stopifnot(nlevels(group) == 2)
  d <- data.frame(dv = as.numeric(dv), group = group)
  d <- d[stats::complete.cases(d), ]
  N <- nrow(d)

  wt <- coin::wilcox_test(dv ~ group, data = d, distribution = "asymptotic")
  Z  <- as.numeric(coin::statistic(wt, type = "standardized"))
  p  <- coin::pvalue(wt)

  w  <- stats::wilcox.test(dv ~ group, data = d, exact = FALSE, correct = FALSE)
  g1 <- d$group == levels(d$group)[1]
  g2 <- d$group == levels(d$group)[2]
  n1 <- sum(g1); n2 <- sum(g2)
  U1 <- as.numeric(w$statistic)
  U2 <- n1 * n2 - U1
  rk <- rank(d$dv)   # pooled ranks -> per-group mean ranks, as SPSS prints them

  list(U1 = U1, U2 = U2, U_min = min(U1, U2), U_max = max(U1, U2),
       Z = Z, absZ = abs(Z), p = p, r = abs(Z) / sqrt(N), N = N, n1 = n1, n2 = n2,
       mean_rank1 = mean(rk[g1]), mean_rank2 = mean(rk[g2]),
       M1 = mean(d$dv[g1]), SD1 = sd(d$dv[g1]),
       M2 = mean(d$dv[g2]), SD2 = sd(d$dv[g2]))
}

# Welch two-sample t test with Cohen's d (pooled SD) --------------------------
# x: 0/1 group code (1 = sacred). t and d are signed sacred - non-sacred (positive
# when the sacred target scores higher).
welch_d <- function(dv, x) {
  d  <- data.frame(dv = as.numeric(dv), g = as.integer(x))
  d  <- d[stats::complete.cases(d), ]
  g1 <- d$dv[d$g == 1]; g0 <- d$dv[d$g == 0]
  tt <- stats::t.test(g1, g0)                                # Welch (unequal variance)
  n1 <- length(g1); n0 <- length(g0)
  sp <- sqrt(((n1 - 1) * stats::var(g1) + (n0 - 1) * stats::var(g0)) / (n1 + n0 - 2))
  tibble(t = unname(tt$statistic), df = unname(tt$parameter), p = tt$p.value,
         d  = (mean(g1) - mean(g0)) / sp,
         M1 = mean(g1), SD1 = stats::sd(g1), M0 = mean(g0), SD0 = stats::sd(g0),
         n1 = n1, n0 = n0)
}

# Simple mediation, Hayes PROCESS Model 4 -------------------------------------
# x: 0/1 (0 = non-sacred, 1 = sacred); m, y: numeric outcomes.
# Indirect effect a*b with a bias-corrected bootstrap CI (the PROCESS default).
mediation_m4 <- function(x, m, y, n_boot = 5000, seed = 2024, conf = 0.95) {
  d <- data.frame(x = as.numeric(x), m = as.numeric(m), y = as.numeric(y))
  d <- d[stats::complete.cases(d), ]

  point <- function(dd) {
    a  <- stats::coef(stats::lm(m ~ x, data = dd))[["x"]]
    bm <- stats::lm(y ~ x + m, data = dd)
    b  <- stats::coef(bm)[["m"]]
    cp <- stats::coef(bm)[["x"]]                              # direct effect c'
    c  <- stats::coef(stats::lm(y ~ x, data = dd))[["x"]]      # total effect c
    c(a = a, b = b, cprime = cp, total = c, indirect = a * b)
  }
  est <- point(d)

  set.seed(seed)                                              # reproducible bootstrap
  n  <- nrow(d)
  bs <- replicate(n_boot, {
    dd <- d[sample.int(n, n, replace = TRUE), ]
    tryCatch(point(dd)[["indirect"]], error = function(e) NA_real_)
  })
  bs <- bs[is.finite(bs)]

  # Bias-corrected interval (BCa with acceleration = 0)
  z0 <- stats::qnorm(mean(bs < est[["indirect"]]))
  za <- stats::qnorm(c((1 - conf) / 2, 1 - (1 - conf) / 2))
  ci <- stats::quantile(bs, c(stats::pnorm(2 * z0 + za[1]),
                              stats::pnorm(2 * z0 + za[2])), names = FALSE)

  list(n = n, n_boot = length(bs), unstd = est, indirect = est[["indirect"]],
       direct = est[["cprime"]], ci = ci, boot_se = stats::sd(bs))
}

# Dual (parallel) mediation with two mediators m1, m2 -------------------------
# Returns the two indirect effects and the direct effect, each with a
# bias-corrected bootstrap CI. Used for the Study 1 selfishness + trust dual model.
mediation_dual <- function(x, m1, m2, y, n_boot = 5000, seed = 2024, conf = 0.95) {
  d <- data.frame(x = as.numeric(x), m1 = as.numeric(m1),
                  m2 = as.numeric(m2), y = as.numeric(y))
  d <- d[stats::complete.cases(d), ]
  point <- function(dd) {
    a1 <- stats::coef(stats::lm(m1 ~ x, data = dd))[["x"]]
    a2 <- stats::coef(stats::lm(m2 ~ x, data = dd))[["x"]]
    ym <- stats::lm(y ~ x + m1 + m2, data = dd)
    c(i1 = a1 * stats::coef(ym)[["m1"]], i2 = a2 * stats::coef(ym)[["m2"]],
      dir = stats::coef(ym)[["x"]])
  }
  est <- point(d)
  set.seed(seed); n <- nrow(d)
  bs <- replicate(n_boot, {
    dd <- d[sample.int(n, n, replace = TRUE), ]
    tryCatch(point(dd), error = function(e) c(i1 = NA, i2 = NA, dir = NA))
  })
  bc <- function(pt, draws) {
    draws <- draws[is.finite(draws)]
    z0 <- stats::qnorm(mean(draws < pt))
    za <- stats::qnorm(c((1 - conf) / 2, 1 - (1 - conf) / 2))
    stats::quantile(draws, stats::pnorm(2 * z0 + za), names = FALSE)
  }
  list(est = est,
       i1 = bc(est[["i1"]], bs["i1", ]),
       i2 = bc(est[["i2"]], bs["i2", ]),
       dir = bc(est[["dir"]], bs["dir", ]))
}

# Probability of superiority / common-language effect size (CLES) -------------
# P(a random draw from `high` exceeds a random draw from `low`); ties count 1/2.
prob_superiority <- function(high, low) {
  high <- high[!is.na(high)]; low <- low[!is.na(low)]
  mean(outer(high, low, function(a, b) (a > b) + 0.5 * (a == b)))
}
```

```{r audit-helpers}
# Build one audit row: computed value vs the paper's reported (or corrected) value.
# tol = absolute tolerance for a "match"; up to 3*tol is "close".
audit_row <- function(quantity, computed, claimed, tol, digits = 2) {
  fmt <- function(v) if (is.na(v)) "--" else formatC(v, format = "f", digits = digits)
  check <- if (is.na(claimed)) "not stated in text"
           else {
             d <- abs(computed - claimed)
             if (d <= tol) "match" else if (d <= 3 * tol) "close" else "DIFFERS"
           }
  tibble(Quantity   = quantity,
         Reported   = fmt(claimed),
         Computed   = fmt(computed),
         Check      = check)
}

# Per-item comparison table: Welch t / d for each item, with the SI-printed t and d
# shown alongside as a reproduction check (computed and SI values should agree).
item_table <- function(df, items, labels, si_t, si_d) {
  dplyr::bind_rows(lapply(seq_along(items), function(i) {
    r <- welch_d(df[[items[i]]], df$x)
    tibble(Item        = labels[i],
           `M sacred`  = sprintf("%.2f", r$M1),
           `M non-sac` = sprintf("%.2f", r$M0),
           t           = sprintf("%.2f", r$t),
           df          = sprintf("%.1f", r$df),
           d           = sprintf("%.2f", r$d),
           `SI t`      = sprintf("%.2f", si_t[i]),
           `SI d`      = sprintf("%.2f", si_d[i]))
  }))
}
```

---

# Study 1 --- Sacred Ring (within-culture)

> *"Alex"* refuses (sacred) or accepts (non-sacred) \$20,000 for a wedding ring bought
> for \$2,000. U.S. representative sample. Reported $N = 281$.

## Load, exclude, and build measures

```{r s1-prep}
s1 <- read_sav(s1_file)

# Pre-registered exclusions: keep Exclusion == 0 ("Good").
s1a <- s1 %>% filter(Exclusion == 0)

# Condition: string "Sell" / "Sacred" -> x = 1 for the sacred (refuser) target.
cond_raw <- as_factor(s1a$Condition)
s1a$x    <- as.integer(cond_raw == "Sacred")
s1a$cond <- factor(s1a$x, levels = c(0, 1),
                   labels = c("Non-sacred (sold)", "Sacred (refused)"))

# Composites and single-item measures (per OSF variable key):
#   trust    = mean(Trust_Att_1..6)  [committed, respectful, trustworthy,
#                                      credible, reliable, benevolent]
#   accept   = mean(Rel_1..4)        [friend, romantic, colleague, boss]
#   selfish  = Trust_Att_7           (NOT part of the trust scale)
#   joint    = Joint_D               (importance of joint decision-making)
#   psv      = P_SV                  (participant's own sacralizing of symbolic items)
s1a$trust   <- composite(s1a, paste0("Trust_Att_", 1:6))
s1a$accept  <- composite(s1a, paste0("Rel_", 1:4))
s1a$selfish <- as.numeric(s1a$Trust_Att_7)
s1a$joint   <- as.numeric(s1a$Joint_D)
s1a$psv     <- as.numeric(s1a$P_SV)

# Push-button check: the data must have the N the paper claims.
stopifnot("Study 1 N after exclusions != 281" = nrow(s1a) == 281L)

# And our hand-built composites must equal the pre-baked OSF columns.
stopifnot(
  max(abs(s1a$trust  - s1a$Trust_C),  na.rm = TRUE) < 1e-6,
  max(abs(s1a$accept - s1a$Rel_Comp), na.rm = TRUE) < 1e-6
)

table(s1a$cond)
```

## Scale reliability

```{r s1-alpha}
s1_alpha_trust  <- cronbach_alpha(s1a, paste0("Trust_Att_", 1:6))
s1_alpha_accept <- cronbach_alpha(s1a, paste0("Rel_", 1:4))
c(trust_alpha = round(s1_alpha_trust, 3), accept_alpha = round(s1_alpha_accept, 3))
```

## Descriptives by condition

```{r s1-desc}
s1_desc <- s1a %>%
  group_by(Condition = cond) %>%
  summarise(n = n(),
            trust_M  = mean(trust),  trust_SD  = sd(trust),
            accept_M = mean(accept), accept_SD = sd(accept),
            .groups = "drop")
s1_desc
```

## Trustworthiness contrast (Mann-Whitney U) --- *main text*

```{r s1-mw}
s1_mw  <- mann_whitney(s1a$trust, s1a$cond)
s1_pos <- prob_superiority(s1a$trust[s1a$x == 1], s1a$trust[s1a$x == 0])

with(s1_mw, cat(sprintf(
  "U_max = %.1f (U_min = %.1f) | Z = %.2f | p = %.2e | r = %.2f | N = %d\n",
  U_max, U_min, absZ, p, r, N)))
cat(sprintf("Probability of superiority / CLES (sacred > non-sacred) = %.3f\n", s1_pos))
```

```{r s1-fig, fig.cap="Study 1: distribution of trustworthiness ratings by condition. Diamonds mark condition means."}
ggplot(s1a, aes(cond, trust, fill = cond)) +
  geom_violin(alpha = 0.5, colour = NA, width = 0.9) +
  geom_jitter(width = 0.08, height = 0, alpha = 0.25, size = 0.9) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 4,
               fill = "white", colour = "black") +
  scale_fill_manual(values = c("#9aa0a6", "#1a73e8"), guide = "none") +
  labs(x = NULL, y = "Trustworthiness (0--100)") +
  theme_minimal(base_size = 13)
```

## Robustness: observers who do not share the value --- *main text*

The effect does not require the observer to hold the particular value. We re-run the
contrast within participants who scored **low** on personal sanctification of symbolic
objects (`P_SV_GRP == 1`, i.e. $\geq 1$ SD below the mean).

```{r s1-subgroup}
# Verify P_SV_GRP grouping is consistent with the continuous P_SV composite
stopifnot(all(s1$P_SV[s1$P_SV_GRP == 1] <= 2.8))
stopifnot(all(s1$P_SV[s1$P_SV_GRP == 3] >= 5.6))

s1_low <- s1a %>% filter(P_SV_GRP == 1)
s1_mw_low <- mann_whitney(s1_low$trust, s1_low$cond)

with(s1_mw_low, cat(sprintf(
  "Low-sanctity subgroup (n = %d): U_max = %.1f | Z = %.2f | p = %.3f | r = %.2f\n",
  N, U_max, absZ, p, r)))
```

## Normality assumption tests --- *SI*

The SI reports Shapiro-Wilk tests to justify the non-parametric main analysis. All four
key variables depart from normality.

```{r s1-normality}
s1_norm <- tibble(
  Variable = c("Trust attribution", "Interpersonal relationships",
               "Joint decision-making", "Participant's own sacred values"),
  W = c(shapiro.test(s1a$trust)$statistic,  shapiro.test(s1a$accept)$statistic,
        shapiro.test(s1a$joint)$statistic,  shapiro.test(s1a$psv)$statistic),
  p = c(shapiro.test(s1a$trust)$p.value,    shapiro.test(s1a$accept)$p.value,
        shapiro.test(s1a$joint)$p.value,    shapiro.test(s1a$psv)$p.value),
  `SI W` = c(.94, .94, .83, .98)
)
s1_norm %>% mutate(W = round(W, 3), p = signif(p, 3))
```

## Pre-registered parametric tests --- *SI*

The SI reports the pre-registered *t*-tests (main analysis uses non-parametric tests
because of the normality violations above). Composite trust and composite interpersonal
relationships first, then trait by trait and relationship by relationship. Cohen's *d*
uses the pooled SD.

```{r s1-parametric}
s1_t_trust  <- welch_d(s1a$trust,  s1a$x)
s1_t_accept <- welch_d(s1a$accept, s1a$x)

cat(sprintf("Trust composite:          t(%.2f) = %.2f, p = %.2e, d = %.2f\n",
            s1_t_trust$df,  s1_t_trust$t,  s1_t_trust$p,  s1_t_trust$d))
cat(sprintf("Interpersonal composite:  t(%.2f) = %.2f, p = %.2e, d = %.2f\n",
            s1_t_accept$df, s1_t_accept$t, s1_t_accept$p, s1_t_accept$d))
```

```{r s1-traits}
# Per-trait t-tests, with the SI-reported t and d alongside for direct comparison.
item_table(
  s1a, paste0("Trust_Att_", 1:6),
  c("Committed", "Respectful", "Trustworthy", "Credible", "Reliable", "Benevolent"),
  si_t = c(16.12, 14.65, 14.20, 11.39, 12.88, 6.62),
  si_d = c(1.90,  1.74,  1.69,  1.35,  1.53,  0.79)
)
```

```{r s1-rels}
# Per-relationship t-tests, SI values alongside.
item_table(
  s1a, paste0("Rel_", 1:4),
  c("Close friend", "Romantic partner", "Work colleague", "Boss/supervisor"),
  si_t = c(12.45, 12.64, 8.67, 10.39),
  si_d = c(1.48,  1.50,  1.03, 1.24)
)
```

## Main analysis without exclusions --- *SI*

The SI re-runs the primary contrasts on the full sample (no exclusions) as a sensitivity
check.

```{r s1-noexcl}
s1f <- s1
s1f$x     <- as.integer(as_factor(s1f$Condition) == "Sacred")
s1f$cond  <- factor(s1f$x, 0:1, c("Non-sacred", "Sacred"))
s1f$trust <- composite(s1f, paste0("Trust_Att_", 1:6))
s1f$accept<- composite(s1f, paste0("Rel_", 1:4))

s1_mw_ne_tr  <- mann_whitney(s1f$trust,  s1f$cond)
s1_mw_ne_acc <- mann_whitney(s1f$accept, s1f$cond)
s1_t_ne_tr   <- welch_d(s1f$trust,  s1f$x)
s1_t_ne_acc  <- welch_d(s1f$accept, s1f$x)

cat(sprintf("Full N = %d\n", nrow(s1f)))
cat(sprintf("MW trust:          U_max = %.1f | Z = %.2f | r = %.2f\n",
            s1_mw_ne_tr$U_max,  s1_mw_ne_tr$absZ,  s1_mw_ne_tr$r))
cat(sprintf("MW interpersonal:  U_max = %.1f | Z = %.2f | r = %.2f\n",
            s1_mw_ne_acc$U_max, s1_mw_ne_acc$absZ, s1_mw_ne_acc$r))
cat(sprintf("t trust:           t(%.2f) = %.2f, d = %.2f\n",
            s1_t_ne_tr$df,  s1_t_ne_tr$t,  s1_t_ne_tr$d))
cat(sprintf("t interpersonal:   t(%.2f) = %.2f, d = %.2f\n",
            s1_t_ne_acc$df, s1_t_ne_acc$t, s1_t_ne_acc$d))
```

## Exploratory analysis of selfishness --- *SI*

Selfishness (`Trust_Att_7`) is not part of the trust scale. The SI reports that observers
attribute *less* selfishness to the sacred valuer, and that selfishness mediates the
condition effect on both trust and interpersonal relationships.

```{r s1-selfish-mw}
s1_mw_self <- mann_whitney(s1a$selfish, s1a$cond)
with(s1_mw_self, cat(sprintf(
  "Selfishness by condition: U_min = %.1f (U_max = %.1f) | Z = %.2f | r = %.2f | N = %d\n",
  U_min, U_max, Z, r * sign(Z), N)))
# Full SI sentence, per condition (SI: sacred M = 9.28, SD = 16.28, mean rank = 87.92,
# n = 138 -- one missing selfishness rating from the 139 retained; non-sacred M = 43.98,
# SD = 27.18, mean rank = 191.60, n = 142).
with(s1_mw_self, cat(sprintf(
  "  sacred:     M = %.2f, SD = %.2f, mean rank = %.2f, n = %d\n  non-sacred: M = %.2f, SD = %.2f, mean rank = %.2f, n = %d\n",
  M2, SD2, mean_rank2, n2, M1, SD1, mean_rank1, n1)))
```

```{r s1-selfish-med}
# Single mediations: condition -> selfishness -> {trust, interpersonal}.
s1_med_self_tr  <- mediation_m4(s1a$x, s1a$selfish, s1a$trust,  n_boot = 5000, seed = 2024)
s1_med_self_acc <- mediation_m4(s1a$x, s1a$selfish, s1a$accept, n_boot = 5000, seed = 2024)

cat(sprintf("cond -> selfishness -> trust:          indirect B = %.2f, 95%% CI [%.2f, %.2f]\n",
            s1_med_self_tr$indirect,  s1_med_self_tr$ci[1],  s1_med_self_tr$ci[2]))
cat(sprintf("cond -> selfishness -> interpersonal:  indirect B = %.2f, 95%% CI [%.2f, %.2f]\n",
            s1_med_self_acc$indirect, s1_med_self_acc$ci[1], s1_med_self_acc$ci[2]))
```

```{r s1-selfish-dual}
# Dual mediation: condition -> {selfishness, trust} -> interpersonal relationships.
s1_dual <- mediation_dual(s1a$x, s1a$selfish, s1a$trust, s1a$accept,
                          n_boot = 5000, seed = 2024)
cat(sprintf("Selfishness indirect:  B = %.2f, 95%% CI [%.2f, %.2f]\n",
            s1_dual$est[["i1"]], s1_dual$i1[1], s1_dual$i1[2]))
cat(sprintf("Trust indirect:        B = %.2f, 95%% CI [%.2f, %.2f]\n",
            s1_dual$est[["i2"]], s1_dual$i2[1], s1_dual$i2[2]))
cat(sprintf("Direct effect:         B = %.2f, 95%% CI [%.2f, %.2f]\n",
            s1_dual$est[["dir"]], s1_dual$dir[1], s1_dual$dir[2]))
```

## Intercorrelations among key variables --- *SI*

```{r s1-intercorr}
s1_vars <- s1a[, c("trust", "accept", "joint", "psv", "selfish")]
colnames(s1_vars) <- c("Trust", "Interpersonal", "Joint dec.", "Own SV", "Selfishness")
round(cor(s1_vars, use = "pairwise.complete.obs"), 2)
```

```{r s1-intercorr-desc}
tibble(Variable = colnames(s1_vars),
       Mean = round(colMeans(s1_vars, na.rm = TRUE), 2),
       SD   = round(apply(s1_vars, 2, sd, na.rm = TRUE), 2))
```

## Audit: computed vs paper (Study 1) {.tabset}

Reported values are declared here as targets, then compared with the numbers computed
above; the per-item tables above display the SI-printed figure next to the computed one.

```{r s1-audit}
s1_cor <- cor(s1_vars, use = "pairwise.complete.obs")
s1_audit <- dplyr::bind_rows(
  # main text
  audit_row("Sacred trust M",               s1_desc$trust_M[2],   86.5, 0.1, 1),
  audit_row("Non-sacred trust M",           s1_desc$trust_M[1],   56.4, 0.1, 1),
  audit_row("Mann-Whitney U (U_max)",       s1_mw$U_max,        17787, 1.0, 1),
  audit_row("Mann-Whitney Z",               s1_mw$absZ,         11.63, 0.1, 2),
  audit_row("Effect size r",                s1_mw$r,             0.69, 0.01, 2),
  audit_row("CLES / prob. superiority",     s1_pos,              0.90, 0.01, 2),
  audit_row("Trust scale alpha",            s1_alpha_trust,      0.93, 0.01),
  audit_row("Acceptance scale alpha",       s1_alpha_accept,     0.91, 0.01),
  audit_row("Low-sanctity subgroup U",      s1_mw_low$U_max,    322.5, 1.0, 1),
  audit_row("Low-sanctity subgroup Z",      s1_mw_low$absZ,      2.84, 0.1, 2),
  audit_row("Low-sanctity subgroup r",      s1_mw_low$r,         0.43, 0.02, 2),
  # SI: parametric composites
  audit_row("t trust composite",            s1_t_trust$t,       15.47, 0.1, 2),
  audit_row("d trust composite",            s1_t_trust$d,        1.84, 0.03, 2),
  audit_row("t interpersonal composite",    s1_t_accept$t,      13.16, 0.1, 2),
  audit_row("d interpersonal composite",    s1_t_accept$d,       1.56, 0.03, 2),
  # SI: without exclusions
  audit_row("No-excl MW trust U",           s1_mw_ne_tr$U_max,  19525, 1.0, 1),
  audit_row("No-excl MW trust Z",           s1_mw_ne_tr$absZ,   10.86, 0.1, 2),
  audit_row("No-excl MW trust r",           s1_mw_ne_tr$r,       0.63, 0.02, 2),
  audit_row("No-excl MW interp. U",         s1_mw_ne_acc$U_max, 18745, 1.0, 1),
  audit_row("No-excl MW interp. Z",         s1_mw_ne_acc$absZ,   9.84, 0.1, 2),
  # SI: selfishness
  audit_row("Selfishness MW U (U_min)",     s1_mw_self$U_min,    2542, 1.0, 1),
  audit_row("Selfishness MW Z",             s1_mw_self$absZ,    10.78, 0.1, 2),
  audit_row("Selfishness MW r",             s1_mw_self$r,        0.64, 0.02, 2),
  audit_row("Selfishness sacred n",         s1_mw_self$n2,        138, 0.1, 0),
  audit_row("Selfishness non-sacred n",     s1_mw_self$n1,        142, 0.1, 0),
  audit_row("Selfishness sacred mean rank", s1_mw_self$mean_rank2, 87.92, 0.01, 2),
  audit_row("Selfishness non-sac mean rank",s1_mw_self$mean_rank1, 191.60, 0.01, 2),
  audit_row("Selfishness sacred M",         s1_mw_self$M2,       9.28, 0.01, 2),
  audit_row("Selfishness sacred SD",        s1_mw_self$SD2,     16.28, 0.01, 2),
  audit_row("Selfishness non-sacred M",     s1_mw_self$M1,      43.98, 0.01, 2),
  audit_row("Selfishness non-sacred SD",    s1_mw_self$SD1,     27.18, 0.01, 2),
  audit_row("Med selfish->trust B",         s1_med_self_tr$indirect,  11.69, 0.5, 2),
  audit_row("Med selfish->interp B",        s1_med_self_acc$indirect, 14.63, 0.5, 2),
  audit_row("Dual: selfishness indirect",   s1_dual$est[["i1"]],  6.24, 0.5, 2),
  audit_row("Dual: trust indirect",         s1_dual$est[["i2"]], 21.42, 0.6, 2),
  audit_row("Dual: direct effect",          s1_dual$est[["dir"]], 5.02, 0.6, 2),
  # SI: intercorrelations
  audit_row("r(trust, interpersonal)",      s1_cor["Trust", "Interpersonal"],  0.79, 0.01, 2),
  audit_row("r(trust, selfishness)",        s1_cor["Trust", "Selfishness"],   -0.68, 0.01, 2)
)
kable(s1_audit, align = "lrrl")
```

---

# Study 2 --- Pork Investment (out-group)

> *"Ahmed,"* a devout Muslim, refuses (sacred) or accepts (non-sacred) a lucrative but
> religiously forbidden investment. Non-Muslim raters only --- an out-group commitment.
> Reported $N = 287$.

## Load, exclude, and build measures

```{r s2-prep}
s2 <- read_sav(s2_file)

# Pre-registered exclusions: keep Exclusions == 0 ("Good").
s2a <- s2 %>% filter(Exclusions == 0)

# Condition: Cond_01  (0 = non-sacred, 1 = sacred) per OSF variable key.
s2a$x    <- as.numeric(s2a$Cond_01)
s2a$cond <- factor(s2a$x, levels = c(0, 1),
                   labels = c("Non-sacred (accepted)", "Sacred (refused)"))

# Composites and single-item measures:
#   trust    = mean(Trust_Att_1..6) == Trust_At
#   accept   = mean(Rel_1..4)       == Int_Per
#   selfish  = Trust_Att_7
#   svfr     = SV_FR    (perceived sacralizing of close friendship)
#   mc       = MC_IOS   (perceived closeness to the Muslim community)
s2a$trust   <- composite(s2a, paste0("Trust_Att_", 1:6))
s2a$accept  <- composite(s2a, paste0("Rel_", 1:4))
s2a$selfish <- as.numeric(s2a$Trust_Att_7)
s2a$svfr    <- as.numeric(s2a$SV_FR)
s2a$mc      <- as.numeric(s2a$MC_IOS)

stopifnot("Study 2 N after exclusions != 287" = nrow(s2a) == 287L)
stopifnot(
  max(abs(s2a$trust  - s2a$Trust_At), na.rm = TRUE) < 1e-6,
  max(abs(s2a$accept - s2a$Int_Per),  na.rm = TRUE) < 1e-6
)

table(s2a$cond)
```

## Scale reliability

```{r s2-alpha}
s2_alpha_trust  <- cronbach_alpha(s2a, paste0("Trust_Att_", 1:6))
s2_alpha_accept <- cronbach_alpha(s2a, paste0("Rel_", 1:4))
c(trust_alpha = round(s2_alpha_trust, 3), accept_alpha = round(s2_alpha_accept, 3))
```

## Descriptives by condition

```{r s2-desc}
s2_desc <- s2a %>%
  group_by(Condition = cond) %>%
  summarise(n = n(),
            trust_M  = mean(trust),  trust_SD  = sd(trust),
            accept_M = mean(accept), accept_SD = sd(accept),
            .groups = "drop")
s2_desc
```

## Trustworthiness contrast (Mann-Whitney U) --- *main text*

```{r s2-mw}
s2_mw  <- mann_whitney(s2a$trust, s2a$cond)
s2_pos <- prob_superiority(s2a$trust[s2a$x == 1], s2a$trust[s2a$x == 0])

with(s2_mw, cat(sprintf(
  "U_max = %.1f (U_min = %.1f) | Z = %.2f | p = %.2e | r = %.2f | N = %d\n",
  U_max, U_min, absZ, p, r, N)))
cat(sprintf("Probability of superiority / CLES (sacred > non-sacred) = %.3f\n", s2_pos))
```

```{r s2-fig, fig.cap="Study 2: distribution of trustworthiness ratings by condition. Diamonds mark condition means."}
ggplot(s2a, aes(cond, trust, fill = cond)) +
  geom_violin(alpha = 0.5, colour = NA, width = 0.9) +
  geom_jitter(width = 0.08, height = 0, alpha = 0.25, size = 0.9) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 4,
               fill = "white", colour = "black") +
  scale_fill_manual(values = c("#9aa0a6", "#1a73e8"), guide = "none") +
  labs(x = NULL, y = "Trustworthiness (0--100)") +
  theme_minimal(base_size = 13)
```

## Normality assumption tests --- *SI*

```{r s2-normality}
s2_norm <- tibble(
  Variable = c("Trust attribution", "Interpersonal relationships",
               "Closeness to Muslim community"),
  W = c(shapiro.test(s2a$trust)$statistic, shapiro.test(s2a$accept)$statistic,
        shapiro.test(s2a$mc)$statistic),
  p = c(shapiro.test(s2a$trust)$p.value,   shapiro.test(s2a$accept)$p.value,
        shapiro.test(s2a$mc)$p.value),
  `SI W` = c(.95, .98, .83)
)
s2_norm %>% mutate(W = round(W, 3), p = signif(p, 3))
```

## Pre-registered parametric tests --- *SI*

```{r s2-parametric}
s2_t_trust  <- welch_d(s2a$trust,  s2a$x)
s2_t_accept <- welch_d(s2a$accept, s2a$x)

cat(sprintf("Trust composite:          t(%.2f) = %.2f, p = %.2e, d = %.2f\n",
            s2_t_trust$df,  s2_t_trust$t,  s2_t_trust$p,  s2_t_trust$d))
cat(sprintf("Interpersonal composite:  t(%.2f) = %.2f, p = %.2e, d = %.2f  (M sacred = %.2f)\n",
            s2_t_accept$df, s2_t_accept$t, s2_t_accept$p, s2_t_accept$d, s2_t_accept$M1))
```

```{r s2-traits}
item_table(
  s2a, paste0("Trust_Att_", 1:6),
  c("Committed", "Respectful", "Trustworthy", "Credible", "Reliable", "Benevolent"),
  si_t = c(15.89, 12.40, 12.69, 12.27, 11.98, 6.29),
  si_d = c(1.87,  1.46,  1.49,  1.44,  1.41,  0.74)
)
```

```{r s2-rels}
item_table(
  s2a, paste0("Rel_", 1:4),
  c("Close friend", "Romantic partner", "Work colleague", "Boss/supervisor"),
  si_t = c(8.46, 2.67, 7.58, 7.79),
  si_d = c(0.99, 0.32, 0.89, 0.92)
)
```

## Main analysis without exclusions --- *SI*

```{r s2-noexcl}
s2f <- s2
s2f$x     <- as.numeric(s2f$Cond_01)
s2f$cond  <- factor(s2f$x, 0:1, c("Non-sacred", "Sacred"))
s2f$trust <- composite(s2f, paste0("Trust_Att_", 1:6))
s2f$accept<- composite(s2f, paste0("Rel_", 1:4))
s2f$svfr  <- as.numeric(s2f$SV_FR)
s2f$mc    <- as.numeric(s2f$MC_IOS)

s2_ne_tr   <- mann_whitney(s2f$trust,  s2f$cond)
s2_ne_acc  <- mann_whitney(s2f$accept, s2f$cond)
s2_ne_svfr <- mann_whitney(s2f$svfr,   s2f$cond)
s2_ne_mc   <- mann_whitney(s2f$mc,     s2f$cond)

cat(sprintf("Full N = %d\n", nrow(s2f)))
cat(sprintf("MW trust:          U_max = %.1f | Z = %.2f | r = %.2f\n", s2_ne_tr$U_max,   s2_ne_tr$absZ,   s2_ne_tr$r))
cat(sprintf("MW interpersonal:  U_max = %.1f | Z = %.2f | r = %.2f\n", s2_ne_acc$U_max,  s2_ne_acc$absZ,  s2_ne_acc$r))
cat(sprintf("MW sacralizing:    U_max = %.1f | Z = %.2f | r = %.2f\n", s2_ne_svfr$U_max, s2_ne_svfr$absZ, s2_ne_svfr$r))
cat(sprintf("MW community:      U_max = %.1f | Z = %.2f | r = %.2f\n", s2_ne_mc$U_max,   s2_ne_mc$absZ,   s2_ne_mc$r))
# Full SI sentence for the interpersonal contrast, per condition (SI: sacred M = 62.13,
# SD = 20.98, mean rank = 186.05, n = 150; non-sacred M = 41.84, SD = 25.22, mean
# rank = 117.40, n = 152).
with(s2_ne_acc, cat(sprintf(
  "  interpersonal sacred:     M = %.2f, SD = %.2f, mean rank = %.2f, n = %d\n  interpersonal non-sacred: M = %.2f, SD = %.2f, mean rank = %.2f, n = %d\n",
  M2, SD2, mean_rank2, n2, M1, SD1, mean_rank1, n1)))
```

## Exploratory analysis of selfishness --- *SI*

```{r s2-selfish}
s2_mw_self <- mann_whitney(s2a$selfish, s2a$cond)
with(s2_mw_self, cat(sprintf(
  "Selfishness by condition: U_min = %.1f (U_max = %.1f) | Z = %.2f | r = %.2f | N = %d\n",
  U_min, U_max, Z, r * sign(Z), N)))

s2_med_self_tr  <- mediation_m4(s2a$x, s2a$selfish, s2a$trust,  n_boot = 5000, seed = 2024)
s2_med_self_acc <- mediation_m4(s2a$x, s2a$selfish, s2a$accept, n_boot = 5000, seed = 2024)
cat(sprintf("cond -> selfishness -> trust:          indirect B = %.2f, 95%% CI [%.2f, %.2f]\n",
            s2_med_self_tr$indirect,  s2_med_self_tr$ci[1],  s2_med_self_tr$ci[2]))
cat(sprintf("cond -> selfishness -> interpersonal:  indirect B = %.2f, 95%% CI [%.2f, %.2f]\n",
            s2_med_self_acc$indirect, s2_med_self_acc$ci[1], s2_med_self_acc$ci[2]))
```

## Sacralizing of friendship and closeness to community --- *SI*

Two exploratory out-group effects. Closeness to the Muslim community is the effect cited
in the main text (participants saw the sacred valuer as *closer* to his community).

```{r s2-svfr-mc}
s2_mw_svfr <- mann_whitney(s2a$svfr, s2a$cond)
s2_mw_mc   <- mann_whitney(s2a$mc,   s2a$cond)

with(s2_mw_svfr, cat(sprintf(
  "Sacralizing friendship: U_max = %.1f | Z = %.2f | r = %.2f  (M sacred = %.2f, non-sacred = %.2f)\n",
  U_max, absZ, r, mean(s2a$svfr[s2a$x == 1], na.rm = TRUE), mean(s2a$svfr[s2a$x == 0], na.rm = TRUE))))
with(s2_mw_mc, cat(sprintf(
  "Community closeness:    U_max = %.1f | Z = %.2f | r = %.2f  (M sacred = %.2f, non-sacred = %.2f)\n",
  U_max, absZ, r, mean(s2a$mc[s2a$x == 1], na.rm = TRUE), mean(s2a$mc[s2a$x == 0], na.rm = TRUE))))
```

## Intercorrelations among key variables --- *SI*

```{r s2-intercorr}
s2_vars <- s2a[, c("trust", "accept", "svfr", "mc", "selfish")]
colnames(s2_vars) <- c("Trust", "Interpersonal", "Sacr. friend", "Muslim comm.", "Selfishness")
round(cor(s2_vars, use = "pairwise.complete.obs"), 2)
```

```{r s2-intercorr-desc}
tibble(Variable = colnames(s2_vars),
       Mean = round(colMeans(s2_vars, na.rm = TRUE), 2),
       SD   = round(apply(s2_vars, 2, sd, na.rm = TRUE), 2))
```

## Audit: computed vs paper (Study 2)

```{r s2-audit}
s2_cor <- cor(s2_vars, use = "pairwise.complete.obs")
s2_audit <- dplyr::bind_rows(
  # main text
  audit_row("Sacred trust M",               s2_desc$trust_M[2],   81.5, 0.1, 1),
  audit_row("Non-sacred trust M",           s2_desc$trust_M[1],   49.3, 0.1, 1),
  audit_row("Mann-Whitney U (U_max)",       s2_mw$U_max,        18177, 1.0, 1),
  audit_row("Mann-Whitney Z",               s2_mw$absZ,         11.22, 0.1, 2),
  audit_row("Effect size r",                s2_mw$r,             0.66, 0.01, 2),
  audit_row("CLES / prob. superiority",     s2_pos,              0.88, 0.01, 2),
  # SI: parametric composites
  audit_row("t trust composite",            s2_t_trust$t,       14.50, 0.1, 2),
  audit_row("d trust composite",            s2_t_trust$d,        1.71, 0.03, 2),
  audit_row("t interpersonal composite",    s2_t_accept$t,       7.85, 0.1, 2),
  audit_row("d interpersonal composite",    s2_t_accept$d,       0.92, 0.03, 2),
  audit_row("Interpersonal M sacred",        s2_t_accept$M1,   62.49, 0.1, 2),
  # SI: without exclusions
  audit_row("No-excl MW trust U",           s2_ne_tr$U_max,   19836.5, 1.0, 1),
  audit_row("No-excl MW trust Z",           s2_ne_tr$absZ,      11.12, 0.1, 2),
  audit_row("No-excl MW interp. U",         s2_ne_acc$U_max,  16582.5, 1.0, 1),
  audit_row("No-excl MW interp. Z",         s2_ne_acc$absZ,      6.83, 0.1, 2),
  audit_row("No-excl MW interp. r",         s2_ne_acc$r,         0.39, 0.02, 2),
  audit_row("No-excl interp. sacred M",     s2_ne_acc$M2,       62.13, 0.01, 2),
  audit_row("No-excl interp. sacred SD",    s2_ne_acc$SD2,      20.98, 0.01, 2),
  audit_row("No-excl interp. sacred rank",  s2_ne_acc$mean_rank2, 186.05, 0.01, 2),
  audit_row("No-excl interp. sacred n",     s2_ne_acc$n2,         150, 0.1, 0),
  audit_row("No-excl interp. non-sac M",    s2_ne_acc$M1,       41.84, 0.01, 2),
  audit_row("No-excl interp. non-sac SD",   s2_ne_acc$SD1,      25.22, 0.01, 2),
  audit_row("No-excl interp. non-sac rank", s2_ne_acc$mean_rank1, 117.40, 0.01, 2),
  audit_row("No-excl interp. non-sac n",    s2_ne_acc$n1,         152, 0.1, 0),
  audit_row("No-excl MW sacralizing U",     s2_ne_svfr$U_max, 17963.5, 1.0, 1),
  audit_row("No-excl MW community U",       s2_ne_mc$U_max,   19832.5, 1.0, 1),
  # SI: selfishness
  audit_row("Selfishness MW U (U_min)",     s2_mw_self$U_min,    3450, 1.0, 1),
  audit_row("Selfishness MW Z",             s2_mw_self$absZ,     9.76, 0.1, 2),
  audit_row("Selfishness MW r",             s2_mw_self$r,        0.58, 0.02, 2),
  audit_row("Med selfish->trust B",         s2_med_self_tr$indirect,  11.75, 0.5, 2),
  audit_row("Med selfish->interp B",        s2_med_self_acc$indirect, 12.07, 0.5, 2),
  # SI: sacralizing / community (excluded sample) + intercorrelations
  audit_row("Sacralizing friend U (U_max)", s2_mw_svfr$U_max, 16455.5, 1.0, 1),
  audit_row("Sacralizing friend r",         s2_mw_svfr$r,        0.52, 0.02, 2),
  audit_row("Community closeness U (U_max)",s2_mw_mc$U_max,   17897.5, 1.0, 1),
  audit_row("Community closeness r",        s2_mw_mc$r,          0.66, 0.02, 2),
  audit_row("r(trust, interpersonal)",      s2_cor["Trust", "Interpersonal"], 0.74, 0.01, 2),
  audit_row("r(trust, Muslim comm.)",       s2_cor["Trust", "Muslim comm."],  0.71, 0.01, 2)
)
kable(s2_audit, align = "lrrl")
```

---

# Audit summary

```{r summary, echo=FALSE}
all_audit <- dplyr::bind_rows(
  dplyr::mutate(s1_audit, Study = "Study 1", .before = 1),
  dplyr::mutate(s2_audit, Study = "Study 2", .before = 1)
)
n_match   <- sum(all_audit$Check == "match")
n_close   <- sum(all_audit$Check == "close")
n_differ  <- sum(all_audit$Check == "DIFFERS")
n_total   <- nrow(all_audit)

cat(sprintf("Audited claims: %d | match: %d | close (within bootstrap/rounding noise): %d | DIFFERS: %d\n",
            n_total, n_match, n_close, n_differ))
if (n_differ > 0) {
  cat("\nClaims that differ from the paper:\n")
  print(knitr::kable(subset(all_audit, Check == "DIFFERS",
                            select = c("Study", "Quantity", "Reported", "Computed"))))
}
```

# Reporting conventions

Every statistic reported for Studies 1 and 2 --- in both the main text and the Supporting
Information --- reproduces from the raw data; the audit summary above records no
discrepancies. Two reporting conventions are worth flagging for a reader comparing this
document line by line with the SI:

- **Mann-Whitney *U*.** The trustworthiness contrasts report *U*~max~ (the larger of the
  two *U* statistics); the exploratory selfishness contrasts report *U*~min~ (2,542 in
  Study 1; 3,450 in Study 2), matching the SI. The standardized *Z* and the effect size
  *r* are convention-independent.
- **Cohen's *d*.** All *d* values use the pooled within-group SD.

---

# Reproducibility record

Everything above ran from the raw `.sav` files with the single seed set at the top. The
software environment is recorded below so a future reader can reproduce the bit-level
results.

```{r session, echo=FALSE}
si <- sessionInfo()
cat("R version:", si$R.version$version.string, "\n")
cat("Platform :", si$platform, "\n")
cat("Seed     : 20260619 (global); 2024 (every mediation bootstrap)\n\n")
cat("Key package versions:\n")
pkgs <- c("haven", "dplyr", "coin", "ggplot2", "tibble", "knitr", "rmarkdown")
for (p in pkgs) {
  v <- tryCatch(as.character(packageVersion(p)), error = function(e) "not installed")
  cat(sprintf("  %-10s %s\n", p, v))
}
```

```{r session-full, echo=FALSE}
sessionInfo()
```
